### 2023-02-23 bg2.Rmd
fix minor bugs

---
title: "Importance of background gene list for enrichment analysis"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/background

## Intro

I previously wrote about the importance of a background gene list for
accurate and correct enrichment analysis.

Here I will demonstrate how important this is, as well as providing a
guide for how to make your own background gene list.

For this guide I will be using data from a previous analysis, which is
deposited at NCBI GEO and SRA under accession numbers: GSE55123 and SRP038101.

The data was presented in a [paper by Lund et al, 2014](https://pubmed.ncbi.nlm.nih.gov/25315154/)

The experiment is designed to investigate the effect of Azacitidine treatment on AML3 cells.

```{r,libs}

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("vioplot")
  library("clusterProfiler")
  library("eulerr")
})

```

## Fetch data

Using RNAseq data from [DEE2](https://pubmed.ncbi.nlm.nih.gov/25315154/),
which can be accessed from dee2.io

```{r,fetch}

mdat <- getDEE2Metadata("hsapiens")

# get sample sheet
ss <- subset(mdat,SRP_accession=="SRP038101")

# fetch the whole set of RNA-seq data
x <- getDEE2("hsapiens",ss$SRR_accession , metadata=mdat, legacy=TRUE)
mx <- x$GeneCounts
rownames(mx) <- paste(rownames(mx),x$GeneInfo$GeneSymbol)
dim(mx)

# aza no filtering
ss$trt <- grepl("Treated",ss$Experiment_title)

ss %>%
  kbl(caption="sample sheet for Aza treatment in AML3 cells") %>%
  kable_paper("hover", full_width = F)

```

## Analysis of read counts

Although I know this data set is good as I was involved in the project,
it is a good idea to show at least the number of reads for each sample.

```{r,counts1}

par(mar=c(5,7,5,1))
barplot(rev(colSums(mx)),horiz=TRUE,las=1,main="number of reads per sample in SRP038101")

```

Now we can have a look at the read count distribution for one of the samples.

```{r,dist1}

# look at the distribution of counts
l <- lapply(1:ncol(mx),function(i) log10( mx[,i] +1 ))
names(l) <- colnames(mx)
l <- rev(l)

blah()
hist(l[[6]],main="Distribution of read counts in SRR1171523",
  xlab="log10(readcounts+1)",breaks=30)

abline(v=1,col="red",lty=2)

```

We can see that most genes have a very low read count, consistent
with most genes being undetectable in any sample.

Based on this typical distribution of read counts, I would consider
any gene with fewer than 10 reads as not detected.
This could be considered a strict detection threshold.
Setting a strict threshold like this could be helpful in selecting
genes that are likely to validate with qRT-PCR.

I will run DESeq2 with and without this threshold to see how
it impacts the results.

## Differential expression analysis with and without filtering

To see whether this has any effect on downstream analysis, DESeq2 will
be run on the same dataset without any filtering, with light filtering and with
strict filtering.

Here I define light filter threshold as the lowest baseline expression level that
DE genes start to appear.
I know this contrast will yield thousands of DE genes, so this is a reasonable
way to set the detection threshold.
The DE genes with lowest baseline expression showed mean read counts of about 2.

```{r,de1}

# DESeq2 without any fitering
dds <- DESeqDataSetFromMatrix(countData = mx , colData = ss, design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <- cbind(as.data.frame(z),mx)
de <- as.data.frame(zz[order(zz$pvalue),])

head(de,10) %>%
  kbl(caption="Top DE genes for Aza treatment (unfiltered)") %>%
  kable_paper("hover", full_width = F)

DET=nrow(mx)
NSIG=nrow(sig)
NUP=nrow(subset(sig,log2FoldChange>0))
NDN=nrow(subset(sig,log2FoldChange<0))

HEADER=paste(DET,"genes included;",NSIG,"w/FDR<0.05;",NUP,"up;",NDN,"down")

plot(log10(de$baseMean) ,de$log2FoldChange,
  cex=0.6,pch=19,col="darkgray",
  main="no detection filter",
  xlab="log10(basemean)",ylab="log2(fold change)")

points(log10(sig$baseMean) ,sig$log2FoldChange,
  cex=0.6,pch=19,col="red")

mtext(HEADER)

abline(v=1,lty=2)

deup <- rownames(subset(de,padj<=0.05 & log2FoldChange>0))
deup <- unique(sapply(strsplit(deup," "),"[[",2))

dedn <- rownames(subset(de,padj<=0.05 & log2FoldChange<0))
dedn <- unique(sapply(strsplit(dedn," "),"[[",2))

all <- rownames(de)
all <- unique(sapply(strsplit(all," "),"[[",2))

```

look at the lowest basemean genes

```{r,low}

sig <- sig[order(sig$baseMean),]

head(sig,10) %>%
  kbl(caption="lowest basemean genes") %>%
  kable_paper("hover", full_width = F)

```

The counts look quite low.
Not sure that these DEGs could be validated with qRT-PCR.

## Now run DESeq2 with filtering

```{r,de2}

mxf <- mx[which(rowMeans(mx)>=10),]
dim(mxf)
dds <- DESeqDataSetFromMatrix(countData = mxf , colData = ss, design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),mxf)
def <-as.data.frame(zz[order(zz$pvalue),])

head(def,10) %>%
  kbl(caption="Top DE genes for Aza treatment (filtered)") %>%
  kable_paper("hover", full_width = F)

sigf <- subset(def,padj<=0.05)

DET=nrow(mxf)
NSIG=nrow(sigf)
NUP=nrow(subset(sigf,log2FoldChange>0))
NDN=nrow(subset(sigf,log2FoldChange<0))

HEADER=paste(DET,"detected genes;",NSIG,"w/FDR<0.05;",NUP,"up;",NDN,"down")

plot(log10(def$baseMean) ,def$log2FoldChange,
  cex=0.6,pch=19,col="darkgray",
  main="strict filter",
  xlab="log10(basemean)",ylab="log2(fold change)")

points(log10(sigf$baseMean) ,sigf$log2FoldChange,
  cex=0.6,pch=19,col="red")

mtext(HEADER)

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",2))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",2))

```

## Now run ORA based enrichment analysis

I will run enrichment analysis using the correct correct background list,
and compare the results to the incorrect whole genome background.

```{r,gs}

if (! file.exists("ReactomePathways.gmt")) {
  download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip", 
    destfile="ReactomePathways.gmt.zip")
  unzip("ReactomePathways.gmt.zip")
}

genesets2 <- read.gmt("ReactomePathways.gmt")

```

### ORA with correct background list.

Firstly with the upregulated genes

```{r,ora2up}

ora2_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora2_up$geneID <- NULL
ora2_up <- subset(ora2_up,p.adjust<0.05 & Count >=10)
ora2_ups <- rownames(ora2_up)

gr <- as.numeric(sapply(strsplit(ora2_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora2_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora2_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora2_up$BgRatio,"/"),"[[",2))

ora2_up$es <- gr/br
ora2_up <- ora2_up[order(-ora2_up$es),]
ora2_up$Description=NULL

head(ora2_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment (filtered)") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(ora2_up$es,20))
names(topup2) <- rev(head(ora2_up$ID,20))

```

Now with the downregulated genes

```{r,ora2dn}

ora2_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora2_dn$geneID <- NULL
ora2_dn <- subset(ora2_dn,p.adjust<0.05 & Count >=10)
ora2_dns <- rownames(ora2_dn)

gr <- as.numeric(sapply(strsplit(ora2_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora2_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora2_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora2_dn$BgRatio,"/"),"[[",2))

ora2_dn$es <- gr/br
ora2_dn <- ora2_dn[order(-ora2_dn$es),]
ora2_dn$Description=NULL

head(ora2_dn) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in Aza treatment (filtered)") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(ora2_dn$es,20)
names(topdn2) <- head(ora2_dn$ID,20)

```

Make a barplot

```{r,topbarplot2}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",20),rep("red",20))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES")

mtext("correct background")

```

### ORA with whole genome background list.

Firstly with the upregulated genes

```{r,ora1up}

ora1_up <- as.data.frame(enricher(gene = defup ,
  universe = all,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora1_up$geneID <- NULL
ora1_up <- subset(ora1_up,p.adjust<0.05 & Count >=10)
ora1_ups <- rownames(ora1_up)

gr <- as.numeric(sapply(strsplit(ora1_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora1_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora1_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora1_up$BgRatio,"/"),"[[",2))

ora1_up$es <- gr/br
ora1_up <- ora1_up[order(-ora1_up$es),]
ora1_up$Description=NULL

head(ora1_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment (unfiltered)") %>%
  kable_paper("hover", full_width = F)

topup1 <- rev(head(ora1_up$es,20))
names(topup1) <- rev(head(ora1_up$ID,20))

```

Now with the downregulated genes

```{r,ora1dn}

ora1_dn <- as.data.frame(enricher(gene = defdn ,
  universe = all,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora1_dn$geneID <- NULL
ora1_dn <- subset(ora1_dn,p.adjust<0.05 & Count >=10)
ora1_dns <- rownames(ora1_dn)

gr <- as.numeric(sapply(strsplit(ora1_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora1_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora1_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora1_dn$BgRatio,"/"),"[[",2))

ora1_dn$es <- gr/br
ora1_dn <- ora1_dn[order(-ora1_dn$es),]
ora1_dn$Description=NULL

head(ora1_dn) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in Aza treatment (unfiltered)") %>%
  kable_paper("hover", full_width = F)

topdn1 <- head(ora1_dn$es,20)
names(topdn1) <- head(ora1_dn$ID,20)

```

Make a barplot

```{r,topbarplot1}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",20),rep("red",20))

barplot(c(topdn1,topup1),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES")

mtext("whole genome background")

```

## Compare pathway results

```{r,venn1,fig.width=6,fig.height=6}

v1 <- list("correct up"=ora2_ups, "wg up"=ora1_ups)

v2 <- list("correct dn"=ora2_dns, "wg dn"=ora1_dns)

par(mar=c(10,10,10,10))

par(mfrow=c(2,1))
plot(euler(v1),quantities = TRUE)
plot(euler(v2),quantities = TRUE)

```

Jaccard index

```{r,jaccard}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

correct <- c(ora2_ups,ora2_dns)
incorrect <- c(ora1_ups,ora1_dns)

jaccard(correct,incorrect)

```

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,session}

sessionInfo()

```

</details>