### 2023-07-23 analysis/example_proteomics.Rmd
starting

---
title: "Examining background gene lists in clusterProfiler - proteomics example"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/enrichment_recipe

## Introduction

This guide is a Rmarkdown script that conducts differential expression and enrichment analysis,
which are very popular workflows for transcriptome data.

The goal of this work is to understand how ClusterProfiler manages the background list,
as we have observed some weird behaviour.
In particular we see that when we provide a custom background, those genes do not appear in the
final analysis unless they are also included in the gene list.

In the code chunk below called `libs`, you can add and remove required R library dependancies.
Check that the libraries listed here match the Dockerfile, otherwise you might get errors.

```{r,libs}

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
})

```

For this report I will be using proteomics data from Mulvey et al 2021 (PMID: 34599159),
who looked at the effect of LPS on THP-1 cell lines.
Although in this paper they looked at numerous timepoints, I will be looking only at the 24hr time
point where the number of significantly differentially expressed proteins was the greatest.
The data were obtained from Supplementary Table 1 and saved to the /data folder.

```{r,fetch}

def <- read.table("../data/lps_24h.tsv",sep="\t",header=TRUE)

```

# KEGG pathway analysis

In the next sections I will run enrichment analysis with over-representation analysis (ORA) test
and compare it to functional class scoring.
I will also investigate some strange behviour of the ORA tool clusterprofiler.

## ORA with Clusterprofiler custom background with otherwise default analysis

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 |

Here I provide a custom background gene list.

Get the gene sets loaded in R.
These are KEGG gene sets

```{r,gs1}
# from https://www.gsea-msigdb.org/gsea/msigdb/human/collections.jsp 16th June 2023
genesets2 <- read.gmt("../ref/c2.cp.kegg.v2023.1.Hs.symbols.gmt")
gsets <- gmtPathways("../ref/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

message(paste("number of genes described in the annotation set:",length(unique(genesets2$gene))))

```

Now filter the gene names into three lists, up-regulated, down-regulated and
background.
Background is simply all genes that were detected.

```{r,subset1}

defup <- unique(subset(def,adj.P.Val<=0.05 & logFC>0)$GN)

defdn <- unique(subset(def,adj.P.Val<=0.05 & logFC<0)$GN)

bg <- unique(def$GN)

message(paste("number of proteins in background:",length(bg)))

```

Enrichment firstly with upregulated genes

```{r,oraup1}

n_pw=10

ora_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_up$geneID <- NULL
ora_up <- subset(ora_up,p.adjust<0.05 & Count >=10)
ora_ups <- rownames(ora_up)

gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

ora_up$es <- gr/br
ora_up <- ora_up[order(-ora_up$es),]
ora_up$Description=NULL

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(ora_up$es,10))
names(topup2) <- rev(head(ora_up$ID,10))

```

The reported background size does not match the dataset.

Now repeat with the downregulated genes

```{r,oradn1}

ora_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_dn$geneID <- NULL
ora_dn <- subset(ora_dn,p.adjust<0.05 & Count >=10)
ora_dns <- rownames(ora_dn)

gr <- as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",2))

ora_dn$es <- gr/br
ora_dn <- ora_dn[order(-ora_dn$es),]
ora_dn$Description=NULL

head(ora_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(ora_dn$es,n_pw)
names(topdn2) <- head(ora_dn$ID,n_pw)

```

Make a barplot

```{r,topbarplot1}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",n_pw),rep("red",n_pw))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE KEGGs",
  xlab="ES",
  cex.main=0.9)

mtext("ORA test")

```

## ORA with Clusterprofiler with custom background and special gene set list to fix the potential bug

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 + BUG FIX |

Here I provide a background gene list for cluterprofiler to use, like a user should.

I have also done some modification to the KEGG gene sets, to en

Now filter the gene names into three lists, up-regulated, down-regulated and
background.
Background is simply all genes that were detected.

Adding all detected genes to the background appears to improve results!

```{r,gs2fix1}

bgdf <- data.frame("background",bg)
colnames(bgdf) <- c("term","gene")
genesets2 <- rbind(genesets2,bgdf)

```

Enrichment firstly with upregulated genes

```{r,oraupfix1}

n_pw=10

orafix_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

orafix_up$geneID <- NULL
orafix_up <- subset(orafix_up,p.adjust<0.05 & Count >=10)
orafix_ups <- rownames(orafix_up)

gr <- as.numeric(sapply(strsplit(orafix_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(orafix_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_up$BgRatio,"/"),"[[",2))

orafix_up$es <- gr/br
orafix_up <- orafix_up[order(-orafix_up$es),]
orafix_up$Description=NULL

head(orafix_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(orafix_up$es,10))
names(topup2) <- rev(head(orafix_up$ID,10))

```

Now with the downregulated genes

```{r,oradnfix1}

orafix_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

orafix_dn$geneID <- NULL
orafix_dn <- subset(orafix_dn,p.adjust<0.05 & Count >=10)
orafix_dns <- rownames(orafix_dn)

gr <- as.numeric(sapply(strsplit(orafix_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(orafix_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_dn$BgRatio,"/"),"[[",2))

orafix_dn$es <- gr/br
orafix_dn <- orafix_dn[order(-orafix_dn$es),]
orafix_dn$Description=NULL

head(orafix_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(orafix_dn$es,n_pw)
names(topdn2) <- head(orafix_dn$ID,n_pw)

```

Make a barplot

```{r,topbarplot2fix1}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",n_pw),rep("red",n_pw))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE KEGGs",
  xlab="ES",
  cex.main=0.9)

mtext("ORA fix test")

```

## Compare pathway results

```{r,venn1,fig.width=8,fig.height=5}

v1 <- list("ORA up"=ora_ups, "ORA fix up"=orafix_ups)
v2 <- list("ORA dn"=ora_dns, "ORA fix dn"=orafix_dns)
v3 <- list("ORA"=union(ora_ups,ora_dns),
  "ORA fix"=union(orafix_ups,orafix_dns))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated KEGG pathways")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated KEGG pathways")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated KEGG pathways")

```

Jaccard index comparing ORA and ORA fix.

```{r,jaccard1}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

ora <- c(ora_ups,ora_dns)
orafix <- c(orafix_ups,orafix_dns)

jaccard(ora,orafix)

```

List gene sets identified only in the "fixed" analysis.

```{r,examine1}

ora <- union(ora_ups,ora_dns)

orafix <- union(orafix_ups,orafix_dns)

setdiff(orafix,ora)

```

# Reactome Gene sets

## ORA with Clusterprofiler custom background with otherwise default analysis

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 |

Here I provide a custom background gene list.

Get the gene sets loaded in R.
These are Reactome gene sets

```{r,gs2}

genesets2 <- read.gmt("../ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("../ref/ReactomePathways_2023-03-06.gmt")

message(paste("number of genes described in the annotation set:",length(unique(genesets2$gene))))

```

Enrichment firstly with upregulated genes

```{r,oraup2}

n_pw=10

ora_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_up$geneID <- NULL
ora_up <- subset(ora_up,p.adjust<0.05 & Count >=10)
ora_ups <- rownames(ora_up)

gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

ora_up$es <- gr/br
ora_up <- ora_up[order(-ora_up$es),]
ora_up$Description=NULL

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(ora_up$es,10))
names(topup2) <- rev(head(ora_up$ID,10))

```

Now repeat with the downregulated genes

```{r,oradn2}

ora_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_dn$geneID <- NULL
ora_dn <- subset(ora_dn,p.adjust<0.05 & Count >=10)
ora_dns <- rownames(ora_dn)

gr <- as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",2))

ora_dn$es <- gr/br
ora_dn <- ora_dn[order(-ora_dn$es),]
ora_dn$Description=NULL

head(ora_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(ora_dn$es,n_pw)
names(topdn2) <- head(ora_dn$ID,n_pw)

```

Make a barplot

```{r,topbarplot22}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",n_pw),rep("red",n_pw))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES",
  cex.main=0.9)

mtext("ORA test")

```

## ORA with Clusterprofiler with custom background and special gene set list to fix the potential bug

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 + BUG FIX |

Adding all detected genes to the background appears to improve results!

```{r,gs2fix2}

bgdf <- data.frame("background",bg)
colnames(bgdf) <- c("term","gene")
genesets2 <- rbind(genesets2,bgdf)

```

Enrichment firstly with upregulated genes

```{r,oraupfix2}

orafix_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

orafix_up$geneID <- NULL
orafix_up <- subset(orafix_up,p.adjust<0.05 & Count >=10)
orafix_ups <- rownames(orafix_up)

gr <- as.numeric(sapply(strsplit(orafix_up$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_up$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(orafix_up$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_up$BgRatio,"/"),"[[",2))

orafix_up$es <- gr/br
orafix_up <- orafix_up[order(-orafix_up$es),]
orafix_up$Description=NULL

head(orafix_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topup2 <- rev(head(orafix_up$es,10))
names(topup2) <- rev(head(orafix_up$ID,10))

```

Now with the downregulated genes

```{r,oradnfix2}

n_pw=10

orafix_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

orafix_dn$geneID <- NULL
orafix_dn <- subset(orafix_dn,p.adjust<0.05 & Count >=10)
orafix_dns <- rownames(orafix_dn)

gr <- as.numeric(sapply(strsplit(orafix_dn$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_dn$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(orafix_dn$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(orafix_dn$BgRatio,"/"),"[[",2))

orafix_dn$es <- gr/br
orafix_dn <- orafix_dn[order(-orafix_dn$es),]
orafix_dn$Description=NULL

head(orafix_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

topdn2 <- head(orafix_dn$es,n_pw)
names(topdn2) <- head(orafix_dn$ID,n_pw)

```

Make a barplot

```{r,topbarplot2fix2}

par(mar=c(5,20,5,1))

cols <- c(rep("blue",n_pw),rep("red",n_pw))

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES",
  cex.main=0.9)

mtext("ORA fix test")

```

## Compare pathway results

```{r,venn2,fig.width=8,fig.height=5}

v1 <- list("ORA up"=ora_ups, "ORA fix up"=orafix_ups)
v2 <- list("ORA dn"=ora_dns, "ORA fix dn"=orafix_dns)
v3 <- list("ORA"=union(ora_ups,ora_dns),
  "ORA fix"=union(orafix_ups,orafix_dns))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated Reactomes")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated Reactomes")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated Reactomes")

```

Jaccard index comparing ORA and ORA fix.

```{r,jaccard2}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

ora <- c(ora_ups,ora_dns)
orafix <- c(orafix_ups,orafix_dns)

jaccard(ora,orafix)

```

List gene sets identified only in the "fixed" analysis.

```{r,examine2}

ora <- union(ora_ups,ora_dns)

orafix <- union(orafix_ups,orafix_dns)

setdiff(orafix,ora)

```

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,session}

sessionInfo()

```

</details>