### 2023-08-02 analysis/example_scrnaseq.Rmd
write gene names out

---
title: "Examining background gene lists in clusterProfiler - scRNA-seq example"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/enrichment_recipe

## Introduction

Load libraries and dataset.

```{r,libs}

load("muscat_vignette.RData")

ls()

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
})

```

```{r,orafunc}

ora <- function(fg,bg,genesets,n_pw=10,mincount=5) {

ora_res <- as.data.frame(enricher(gene = fg ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_res$geneID <- NULL
ora_res <- subset(ora_res,p.adjust<0.05 & Count >=mincount)
ora_res_sets <- rownames(ora_res)

gr <- as.numeric(sapply(strsplit(ora_res$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_res$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_res$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_res$BgRatio,"/"),"[[",2))

ora_res$es <- gr/br
ora_res <- ora_res[order(-ora_res$es),]
ora_res$Description=NULL
return(ora_res)
}

orafix <- function(fg,bg,genesets,n_pw=10,mincount=5) {

bgdf <- data.frame("background",bg)
colnames(bgdf) <- c("term","gene")
genesets_fix <- rbind(genesets,bgdf)

ora_res <- as.data.frame(enricher(gene = fg ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets_fix,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

ora_res$geneID <- NULL
ora_res <- subset(ora_res,p.adjust<0.05 & Count >=mincount)
ora_res_sets <- rownames(ora_res)

gr <- as.numeric(sapply(strsplit(ora_res$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_res$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora_res$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora_res$BgRatio,"/"),"[[",2))

ora_res$es <- gr/br
ora_res <- ora_res[order(-ora_res$es),]
ora_res$Description=NULL
return(ora_res)
}

```

For this report I will use a SingleCellExperiment (SCE) containing 10x droplet-based
scRNA-seq PBCM data from 8 Lupus patients obtained before and after 6h-treatment with IFN-β
 (Kang et al. 2018).

The complete raw data, as well as gene and cell metadata is available through the NCBI GEO,
accession number GSE96583.

Downstream analysis was done according to the Muscat vignette.

Focusing on CD4 T-cells.

```{r,fetch}

names(tbl)

message("Total detected genes in each cluster and differentially expressed")

lapply(tbl, function(x) {
  nsig <- nrow(subset(x,p_adj.loc<0.05))
  ntot <- nrow(x)
  c(ntot,nsig)
} )

head(tbl[["CD4 T cells"]])  %>%
  kbl(caption="Example of muscat dataset. Control vs IFN-beta treatment") %>%
  kable_paper("hover", full_width = F)

```

# KEGG pathway analysis

In the next sections I will run enrichment analysis with over-representation analysis (ORA) test
and compare it to functional class scoring.
I will also investigate some strange behviour of the ORA tool clusterprofiler.

## ORA with Clusterprofiler custom background with otherwise default analysis

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | Muscat FDR<0.05 CD4 T-cells |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | R/Experiment hub "Kang" or GSE96583. |
| Other parameters | Min gene set size of 5 |

Here I provide a custom background gene list.

Get the gene sets loaded in R.
These are KEGG gene sets

```{r,gs1}
# from https://www.gsea-msigdb.org/gsea/msigdb/human/collections.jsp 16th June 2023
genesets2 <- read.gmt("../ref/c2.cp.kegg.v2023.1.Hs.symbols.gmt")
gsets <- gmtPathways("../ref/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

message(paste("number of genes described in the annotation set:",length(unique(genesets2$gene))))

```

Now filter the gene names into three lists, up-regulated, down-regulated and
background.
Background is simply all genes that were detected.

```{r,subset1}

def <- lapply(tbl , function(x) {
  bg <- unique(x$gene)
  up <- unique(subset(x,p_adj.loc<0.05&logFC>0)$gene)
  dn <- unique(subset(x,p_adj.loc<0.05&logFC<0)$gene)
  return(list("bg"=bg,"up"=up,"dn"=dn))
} )

str(def)

```

Enrichment firstly with upregulated genes

```{r,oraup1}

ups <- lapply(1:length(def),function(i) {
  ora(def[[i]]$up,def[[i]]$bg,genesets=genesets2)
})

up_sets <- unlist(lapply(1:length(ups) , function(i) {
  nm <- names(def)[[i]]
  sets <- ups[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(up_sets)
up_sets

```

Now repeat with the downregulated genes

```{r,oradn1}

dns <- lapply(1:length(def),function(i) {
  ora(def[[i]]$dn,def[[i]]$bg,genesets=genesets2)
})

dn_sets <- unlist(lapply(1:length(dns) , function(i) {
  nm <- names(def)[[i]]
  sets <- dns[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(dn_sets)
dn_sets

```

## ORA with Clusterprofiler with custom background and special gene set list to fix the potential bug

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | R/Experiment hub "Kang" or GSE96583. |
| Other parameters | Min gene set size of 10 + BUG FIX |

Here I provide a background gene list for cluterprofiler to use, like a user should.

I have also done some modification to the KEGG gene sets, to en

Now filter the gene names into three lists, up-regulated, down-regulated and
background.
Background is simply all genes that were detected.

Enrichment firstly with upregulated genes

```{r,orafixup1}

fixups <- lapply(1:length(def),function(i) {
  orafix(def[[i]]$up,def[[i]]$bg,genesets=genesets2)
})

fixup_sets <- unlist(lapply(1:length(fixups) , function(i) {
  nm <- names(def)[[i]]
  sets <- fixups[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(fixup_sets)
fixup_sets

```

Now repeat with the downregulated genes

```{r,orafixdn1}

fixdns <- lapply(1:length(def),function(i) {
  orafix(def[[i]]$dn,def[[i]]$bg,genesets=genesets2)
})

fixdn_sets <- unlist(lapply(1:length(fixdns) , function(i) {
  nm <- names(def)[[i]]
  sets <- fixdns[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(fixdn_sets)
fixdn_sets

```

## Compare pathway results

```{r,venn1,fig.width=8,fig.height=5}

v1 <- list("ORA up"=up_sets, "ORA fix up"=fixup_sets)
v2 <- list("ORA dn"=dn_sets, "ORA fix dn"=fixdn_sets)
v3 <- list("ORA"=union(up_sets,dn_sets),
  "ORA fix"=union(fixup_sets,fixdn_sets))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated KEGG pathways")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated KEGG pathways")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated KEGG pathways")

```

Jaccard index comparing ORA and ORA fix.

```{r,jaccard1}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

ora <- c(up_sets,dn_sets)
orafix <- c(fixup_sets,fixdn_sets)

jaccard(ora,orafix)

```

# Reactome Gene sets

## ORA with Clusterprofiler custom background with otherwise default analysis

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | R/Experiment hub "Kang" or GSE96583. |
| Other parameters | Min gene set size of 10 |

Here I provide a custom background gene list.

Get the gene sets loaded in R.
These are Reactome gene sets

```{r,gs2}

genesets2 <- read.gmt("../ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("../ref/ReactomePathways_2023-03-06.gmt")

message(paste("number of genes described in the annotation set:",length(unique(genesets2$gene))))

```

Enrichment firstly with upregulated genes

```{r,oraup2}

ups <- lapply(1:length(def),function(i) {
  ora(def[[i]]$up,def[[i]]$bg,genesets=genesets2)
})

up_sets <- unlist(lapply(1:length(ups) , function(i) {
  nm <- names(def)[[i]]
  sets <- ups[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(up_sets)
up_sets

```

Now repeat with the downregulated genes

```{r,oradn2}

dns <- lapply(1:length(def),function(i) {
  ora(def[[i]]$dn,def[[i]]$bg,genesets=genesets2)
})

dn_sets <- unlist(lapply(1:length(dns) , function(i) {
  nm <- names(def)[[i]]
  sets <- dns[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(dn_sets)
dn_sets

```

## ORA with Clusterprofiler with custom background and special gene set list to fix the potential bug

I've compiled a reporting checklist:

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | R/Experiment hub "Kang" or GSE96583. |
| Other parameters | Min gene set size of 10 + BUG FIX |

```{r,orafixup2}

fixups <- lapply(1:length(def),function(i) {
  orafix(def[[i]]$up,def[[i]]$bg,genesets=genesets2)
})

fixup_sets <- unlist(lapply(1:length(fixups) , function(i) {
  nm <- names(def)[[i]]
  sets <- fixups[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(fixup_sets)
fixup_sets

```

Now repeat with the downregulated genes

```{r,orafixdn2}

fixdns <- lapply(1:length(def),function(i) {
  orafix(def[[i]]$dn,def[[i]]$bg,genesets=genesets2)
})

fixdn_sets <- unlist(lapply(1:length(fixdns) , function(i) {
  nm <- names(def)[[i]]
  sets <- fixdns[[i]]$ID
  if ( length(sets) >0 ) { paste(nm,sets) }
} ) )

length(fixdn_sets)
fixdn_sets

```

## Compare pathway results

```{r,venn2,fig.width=8,fig.height=5}

v1 <- list("ORA up"=up_sets, "ORA fix up"=fixup_sets)
v2 <- list("ORA dn"=dn_sets, "ORA fix dn"=fixdn_sets)
v3 <- list("ORA"=union(up_sets,dn_sets),
  "ORA fix"=union(fixup_sets,fixdn_sets))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated Reactomes")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated Reactomes")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated Reactomes")

```

Jaccard index comparing ORA and ORA fix.

```{r,jaccard2}

jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}

ora <- c(up_sets,dn_sets)
orafix <- c(fixup_sets,fixdn_sets)

jaccard(ora,orafix)

```

List gene sets identified only in the "fixed" analysis.

```{r,examine2}

ora <- union(ora_ups,ora_dns)

orafix <- union(orafix_ups,orafix_dns)

setdiff(orafix,ora)

```

## Session information

For reproducibility

<br><details><summary><b>Click HERE to show session info</b></summary><br><p>

```{r,session}

sessionInfo()

```

</details>