### 2024-05-31 old/manuscript/manuscript.Rmd

---
title: "Importance of using the correct background list for over-representation analysis"
author: "Mark Ziemann, Barry Schroeter, Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: plos-computational-biology.csl
---

Alternate title: *Problems in over-representation analysis software*

[Mark Ziemann<sup>1</sup>*](https://orcid.org/0000-0002-7688-6974),
[Barry Schroeter<sup>1</sup>]()
[Anusuiya Bora<sup>1,2</sup>](https://orcid.org/0009-0006-2908-1352)

**Affiliations**

1. Deakin University, Geelong, Australia, School of Life and Environmental Sciences.

2. Vellore Institute of Technology, Vellore, India.

(*) Corresponding author: m.ziemann@deakin.edu.au

## Abstract

* Over-representation analysis (ORA) is the most common form of functional enrichment and one of the
most used techniques in computational biology, yet there are common methodological issues that
undermine its validity.

* These problems are typically ascribed to errors in analysis and reporting by the end-users,
however software design is another contributor.

* Here, we identify and characterise a potential bug popular ORA software
package and web tool.

* We show that this bug deleteriously affects the results of ORA tests, and we present a simple
work-around to circumvent it.

## Introduction

```{r,libs,echo=FALSE}

library("DiagrammeR")

```

Functional enrichment analysis is a family of computational approaches used to interpret large omics
data.
The most commonly used method is over-representation analysis (ORA) which involves the selection of
genes of interest, followed by a statistical test to evaluate whether genes of certain annotated
categories are over-represented in the set of interest.
The set of interest may be defined in different ways, such as genes that meet a differential expression
significance threshold, or those that have been clustered into a group based on co-expression patterns.
There are various sources of annotated categories, including gene ontologies, protein-coding features,
gene co-expression and signalling or metabolic pathways.
Providers of such annotations include but are not limited to the gene ontology consortium, Reactome,
KEGG, WikiPathways and MSigDB.

Central to ORA tests is the selection of a list of background genes.
This is also sometimes called the "universe" or the "reference" list.
These are defined as any gene that is detected robustly, and has a chance at making the set of interest.
For RNA-seq studies, this could be a threshold detection limit based either on raw counts or counts per
million.

The first use of ORA-based functional enrichment that we are aware of is a 1999 report that used the
hypergeometric test to determine the chance probability of observing over-representation of certain
annotated functions in co-regulated genes in yeast [@Tavazoie1999-kf].
This general approach is the basis of most ORA-based significance testing and can be represented by the
following equation:

$$
P=1-\sum_{i=0}^{k-1}\frac{ {f \choose i} {g-f \choose n-1} }{g \choose n}
$$

Where (P) is the probability of observing at least (k) genes from a functional category within the
list of selected genes with size (n).
Here, (g) is the total number of detected genes, and (f) is the number of detected genes belonging to a
functional category.

The fold enrichment score (ES) is commonly used to express the degree of enrichment and is a ratio-of-ratios.
The foreground ratio is the number of selected genes belonging a functional class (A) divided by the
number of detected genes in the functional class (B).
The background ratio is the number of selected genes not belonging to the functional class (C) divided
by the number of detected genes not in the functional class (D).

$$
ES=\frac{ A / B}{ C / D}
$$

Both equations illustrate that correct definition of the number of background genes is important to
obtain accurate enrichment scores and significance values.
A common error in ORA of differential gene expression is to use all genes in the genome as the
background @Wijesooriya2022-li.
This leads to a severe over-estimation of the number of background genes as there are ~45k annotated
genes in the human genome (according to Ensembl v109), but RNA-seq robustly detects only ~15k genes.
This leads to inflation of p-values and results that share only ~40% similarity with properly conducted
ORA analysis.
This issue is even more severe for other omics, where the number of detected genes is lower, for example
proteomics and single cell RNA-seq where there are only a few thousand detected analytes.
To address this issue, most major ORA tools allow for a user-provided background list of detected genes.

While testing some of these tools, we observed an unusual behaviour whereby the number of background
genes reported by the tool varied significantly from the length of the gene list we provided.
Moreover, this number also varied when we used a different library of gene function annotations.
Puzzled by this behaviour, we browsed the computational biology forums and were met with other confused
users wondering why the number of genes in the background list would vary like this.
The tool developers kindly clarified the situation, by stating that the tool removed genes from the
entire analysis if they were not members of a functional category.
We thought this behaviour odd, because it has the potential to significantly alter the test being
conducted, away from it's original application.

In this paper, we examine the effect of this filtering step on ORA of real RNA-seq data, use simulated
RNA-seq data to determine the impact upon precision and recall, and provide practical recommendations
to end users to obtain the most accurate results.

In a hypothetical example, there were 134 genes classified as detected.
Of these, 38 were defined as upregulated in response to a stimulus based on a differential expression
test.
From the 38, 8 genes belonged to the "cytokine signaling pathway" category.
This category contains 26 genes in total, but only 18 were detected in this study.

So the background can be defined as 18/134 and the foreground as 8/38 and the enrichment score
expressed as (8/38)/(18/135)=1.58

In this example:

*k*=8

*n*=26

*g*=134

*f*=18

We can calculate the probability of getting exactly 7 members of the set.

$$
P=\frac{ {18 \choose 7} {134-18 \choose 26-1} }{134 \choose 26}
$$

$$
P=\frac{ 480700 * 1.71e29 }{ 3.84e+33 } = 1/21.4 = 0.047
$$

This needs to be repeated for other values of *i*, namely 6 to 0 and then sum all of these.

```{r,hyperdist}
res<-unlist(lapply(0:38, function(i) {  (choose(18,i) * choose((134-18),(38-1)) / choose(134,38)) }))
plot(res,type="b")
```

## Methods

### Demonstrating the impact of the problem

To demonstrate the impact of this problem on real omics data, we conducted enrichment analysis of
four different types of data.
We obtained differential proteomics data from Stukalov et al (https://pubmed.ncbi.nlm.nih.gov/33845483/),
who profiled the effect of SARS-COV2 infection on A549 cells.
At 24 hrs infection with SARS-CoV2, Stukalov et al reported the detection of 5825 proteins, including
165 downregulated and 44 upregulated (FDR<0.05).
For bulk RNA-seq, we obtained count data from DEE2.io (Ziemann et al, 2019) for control and azacitidine
exposed AML3 cells (SRA accession SRP038101) (Lund et al, 2014, PMID:25315154).
Genes with fewer than 10 reads per sample on average were considered not detected and were omitted from
downstream analysis.
DESeq2 was used to compare treatment groups and genes with FDR<0.05 were selected (Love 2014).
For single cell RNA-seq 10X Genomics droplet-based, scRNA-seq PBMC data from 8 Lupus patients obtained
before and after 6h-treatment with IFN-β (Kang et al. 2018).
The complete raw data, as well as gene and cell metadata is available through the NCBI GEO,
accession number GSE96583.
Analysis of this data set  was done using a pseudobulk approach according to the Muscat package vignette
(Crowell et al, 2020), yielding lists of detected genes for each cell state as well as up- and
down-regulated gene lists (FDR<0.05).
Infinium 450K BeadChip based Epigenenome Wise Association Study (EWAS) data was obtained from
Joehanes et al (PMID:27651444), a meta-analysis investigating the DNA methylation differences between
smokers and non-smokers.
Data was downloaded from EWAS Catalog at the link (http://ewascatalog.org/?study=27651444_smoking_current_vs_never_smoking).
As a background gene list, all genes represented by probes on the 450K array were listed.
For all of the above analyses, up- and down-regulated genes were partitioned into separate lists,
for separate enrichment analysis.
These gene lists were loaded into R and were subject to ClusterProfiler analysis along with the
respective background gene set.
KEGG (REF) or Reactome (REF) pathway annotations were used (GMT format).
KEGG pathway annotations v2023.1 were obtained from MSigDB (REF), while Reactome pathways were downloaded
from the Reactome website (2023-03-06).
Pathways (gene sets) with fewer than 10 detected genes were not included in the analysis.
In addition to the default ClusterProfiler analysis, a modification was made to circumvent the
background filtering problem.
To do this, the list of detected genes was appended as a new "pathway" to the KEGG or Reactome annotation
set.
The results of this "fixed" analysis was compared those using the default algorithm using an Euler
diagram.

### Examining the breadth of the problem

Next, we investigated the behaviour of many different ORA tools to find out whether they suffered from
the same problem or not.
We analysed the 20 most popular ORA tools as described by Xie et al, 2021 (PMID:33858350).
To understand whether the background list filtering problem was apparent, we used the gene lists from
the bulk RNA-seq dataset, with various pathway /gene set databases.
If the tool reported the number of genes in the background list, we used this as evidence to classify
the tool's behaviour.
If not, then we undertook forensic investigation of the enrichment results to infer the program's
behaviour.

## Results

### Background list filtering biases ORA results

### Inappropriate background list filtering is widespread

## Acknowledgements

This research was supported by use of the Nectar Research Cloud, a collaborative Australian research
platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).

## Availability of materials

* Code repository:

* Example Docker image:

## Bibliography