### 2024-06-15 manuscript/manuscript.Rmd
spellcheck

---
title: "Two subtle problems with over-representation analysis"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: plos-computational-biology.csl
---

[Mark Ziemann<sup>1,2</sup>*](https://orcid.org/0000-0002-7688-6974),
[Anusuiya Bora<sup>1,2</sup>](https://orcid.org/0009-0006-2908-1352)

**Affiliations**

1. Burnet Institute, Melbourne, Australia.

2. Deakin University, School of Life and Environmental Sciences, Geelong, Australia.

(*) Corresponding author: mark.ziemann@burnet.edu.au

# Abstract

Over-representation analysis (ORA) is used widely to assess the enrichment of functional
categories in a gene list compared to a background list.
ORA is therefore a critical method in the interpretation of 'omics data, relating gene lists
to biological functions and themes.
Although ORA is hugely popular, we and others have noticed two potentially undesired
behaviours of some ORA tools.
The first one we call the "background problem", because it involves the software eliminating
large numbers of genes from the background list if they are not annotated as belonging to any
category.
The second one we call the "false discovery rate problem", because some tools underestimate
the true number of parallel tests conducted.
Here we demonstrate the impact of these issues on several real RNA-seq datasets and use
simulated RNA-seq data to quantify the impact of these problems.
We show that the severity of these problems depends on the gene set library, the number of
genes in the list, and the degree of noise in the dataset.
These problems can be mitigated by changing packages/websites for ORA or by changing to another
approach such as functional class scoring.

# Keywords

* Bioinformatics

* Pathway analysis

* Gene set enrichment analysis

* Research integrity

# Main Body

## Introduction

```{r,libs,echo=FALSE}

library("kableExtra")

```

Over-representation analysis (ORA) is a type of functional enrichment analysis (FEA) that
involves the summarisation of omics data to reflect biological differences.
FEA has become one of the most popular methods in bioinformatics,
collectively accumulating 131,332 citations as of late 2019 [@Xie2021-nv].

ORA involves the selection of genes of interest, followed by a test to ascertain whether
certain functional categories are over-represented in the selected genes.
Since its initial development in 1999 [@Tavazoie1999-bu], ORA has proliferated widely,
becoming part of many bioinformatics websites and software packages.
The most popular ORA website is DAVID [@Dennis2003-ar;@Huang2009-gi;@Sherman2022-jn],
followed by Ingenuity Pathway Analysis, a commercial software package provided by QIAGEN Inc.
ORA has been implemented into R/Bioconductor packages including `clusterProfiler`
[@Yu2012-jj;@Wu2021-wy], `limma` [@Ritchie2015-oz] and `GOseq` [@Young2010-iw].

Another popular approach to FEA is functional class scoring (FCS).
In FCS, all detected genes are ranked by their degree of differential expression followed
by a statistical test for enrichment of gene categories at either extreme of the
ranked list [@Subramanian2005-no].

According to a study that used simulated differential expression data, FCS accuracy was
superior to ORA over a range of experimental designs [@Kaspi2020-rn].
Despite this performance difference, a survey of articles published in 2019 showed that
seven of the top eight most popular tools were based on ORA [@Wijesooriya2022-fw].
The discrepancy between performance and popularity is likely due to the relative ease of
conducting ORA as compared to FCS.
A minimal ORA might involve pasting a list of gene symbols into a text box on a website,
with results appearing nearly instantly.
On the other hand, FCS typically involves installing specific software and dealing with a
dataset representing every detected gene (~20,000 rows) and ensuring that the file
format from upstream tools is compatible with FCS packages.

Despite the popularity of ORA, there are concerns that this technique is being misused.
Due to technological and biological reasons, not all genes are detected in transcriptomic
studies, meaning that some genes will more readily appear as differentially expressed.
To address this bias, a background list of detected genes is required for comparison to the
list of selected genes (foreground list) [@Tilford2009-vx].
Failure to use a background gene list leads to dramatic changes in enrichment results,
rendering them unreliable [@Timmons2015-ki;@Wijesooriya2022-fw].
Unfortunately, use of an appropriate background list is reported in only a small fraction
(~4%) of peer-reviewed articles describing enrichment analysis [@Wijesooriya2022-fw].
Correction of p-values for multiple testing is also crucial in controlling the false positive
rate as enrichment analysis typically involves thousands of parallel tests [@Tilford2009-vx].
But appropriate p-value correction is reported in only ~54% of studies [@Wijesooriya2022-fw].
These issues have resulted in the development of best practices for end users and
stronger reporting standards [@Zhao2023-mx].

The focus of this work is to raise awareness to two existing problems in ORA implementations
of some popular enrichment tools.
The first problem is that genes belonging to the background list are removed from the analysis
entirely if they are not annotated as belonging to any functional categories.
This results in the background list appearing smaller than it should, leading to an
underestimation of fold enrichment scores and significance values.
In principle, this problem would differentially affect analyses involving different gene set
libraries, with smaller libraries such as Kyoto Encyclopedia of Genes and Genomes (KEGG)
[@Kanehisa2023-ni], which describes 2,727 genes affected more severely as compared to larger
libraries like Gene Ontology [@The_Gene_Ontology_Consortium2023-xp] which describes 19,428
genes (figures obtained from MSigDB June 2024 [@Liberzon2015-cs]).

The second problem is that adjusting p-values for multiple testing, also known as
false-discovery rate correction, is implemented improperly.
Determining which pathways should be subject to false discovery rate correction depends on
how a detection threshold is set.
Ideally, a pathway should be considered detected based on the presence of a predetermined
number of member genes in the background set.
However, some tools use the foreground set to define a pathway as detected.
This is problematic, as the act of trying to find an intersection between a foreground gene list
and a pathway is a test, even if there are no common genes.
This results in pathways being discarded from the analysis if no common genes are found in
the foreground, despite many genes being present in the background list.
This artificially reduces the number of tests conducted and makes FDR values appear lower
than they should, potentially raising the rate of false positives.
We expect this problem to have a more severe effect when the foreground list is small.

Here, we define the effect of these two problems on real RNA-seq based enrichment analysis
results and demonstrate how these two issues impact performance using simulated expression
data.
Finally, we classify popular tools regarding these two problems and provide general
recommendations for end users.

## Methods

### Expression data preparation

To test the effect of these two problems on RNA-seq based enrichment analysis, data from seven
transcriptomic studies were downloaded from DEE2 using the getDEE2 R/Bioconductor package
[@Ziemann2019-ux].
Raw data for these seven studies are available from NCBI Sequence Read Archive under the
accession numbers SRP128998, SRP038101, SRP037718, SRP096177, SRP097759, SRP253951 and
SRP068733 [@Felisbino2021-hc; @Lund2014-ds; @Rafehi2014-rv; @Keating2014-ab;
@Blanco-Melo2020-hh; @Rafehi2017-qn].
For each dataset, kallisto (v0.43.1) [@Bray2016-lu] transcript counts were aggregated to the
gene level.
Genes with fewer than 10 reads per sample on average were removed from downstream analysis.
Remaining genes passing this selection were included in the background gene set.
Differential expression analysis was conducted with DESeq2 v1.44.0 [@Love2014-wg], with
genes identified by their Ensembl identifiers.
Gene symbols were then fetched using biomaRt v2.60.0, based on Ensembl version 112 [@Durinck2009-dr].

### Examining the background problem using real transcriptome data

Human gene sets were obtained from MSigDB v2023.2 [@Liberzon2015-cs], and included
KEGG Medicus [@Kanehisa2023-ni],
Reactome [@Milacic2024-iv],
Wikipathways [@Agrawal2024-lr],
MicroRNA targets from miRdb [@Chen2020-wo],
transcription factor targets from GTRD [@Kolmykov2021-mn],
Gene Ontology (GO) [@The_Gene_Ontology_Consortium2023-xp],
Human Phenotype Ontology (HPO) [@Kohler2021-ha],
Cell markers and
Hallmark pathways [@Liberzon2015-cs].
To demonstrate the background problem, differentially expressed genes with
FDR<0.05 were subset, with separate lists for up and downregulated genes.
If fewer than 200 genes met this criterion, then the 200 genes with the smallest p-values
were selected, as suggested by a previous report [@Tarca2013-xh].
These gene lists were subjected to ORA using clusterProfiler's `enricher` function
v4.12.0, using a minimum set size of 5 and a no maximum set size.

To systematically assess the impact of the background problem, we conducted ORA with
clusterProfiler using the different gene set libraries described above for the seven RNA-seq
datasets.
ClusterProfiler was used with the same parameters as above, and we devised a simple workaround
to the background problem, which is to append the entire background list as a gene set to the
library.
This forces clusterProfiler to retain all background genes.
The ORA analyses were then filtered for significant sets using an FDR threshold of 0.05.
Jaccard index was used to compare original and corrected analyses.

### Examining the FDR problem

For this test, 2000 of the top up and down regulated genes were selected from each
differential expression dataset based on p-value.
A workaround for the FDR problem was devised by filtering the gene sets by the presence of
at least two genes in the background list prior to running clusterProfiler.
After clusterProfiler using the parameters above, the number of gene sets in the results was
quantified, and compared to the number that should have been reported using the two gene
limit.
The difference between these numbers (*n*) represents the number of missing gene sets.
To account for these tests, *n* values of 1 were appended to the p-values obtained by
clusterProfiler, followed by FDR correction in R, to obtain the properly corrected FDR values.
As above, analysis of seven datasets was done with nine gene set libraries, the significant
sets were selected and the Jaccard statistics were collected for each run to compare original
and corrected analysis.

As we predicted that this effect could be more severe for smaller foreground lists, we
performed parallel analysis with foreground gene lists with sizes varying between 125 and
2,000 genes, as ranked by FDR value.

### Quantifying the effect of these problems on precision and recall using simulated expression data

RNA-seq counts for SRA run accession ERR2539161 were obtained from DEE2 [@Mo2018-zb].
This dataset was selected due to its high number of assigned reads (367M).
A library of 200 random gene sets were generated, each containing 30 genes.
To generate differential expression profiles, 5% of gene sets were selected to be
differentially expressed, with equal numbers of up- and down-regulated pathways.
Pseudosamples were generated by first downsampling the counts using the `thincounts`
function of the `edgeR` package v4.2.0 [@Robinson2010-eb] to 20M reads, followed by the
addition of some extra variation drawn from a normal distribution with mean of 1 and
standard deviation varying between 0.1 and 0.6, this value selected randomly for each gene.
Three "control" and five "case" pseudosamples were generated.
For the genes selected to be differentially expressed, expression values were multiplied by a
log fold change in the case samples of 0.5 for upregulated genes and -0.5 for downregulated
genes.
Genes that were selected to be both up and downregulated were left unchanged.
With these expression profiles, DESeq2 was used for differential analysis, and the results
were subjected to clusterProfiler which suffers from both problems,
clusterProfiler with the background problem fix, clusterProfiler with the FDR problem fix,
clusterProfiler with the fixes to background and FDR problems, an ORA function called `fora`
from `fgsea` bioconductor package v1.30.0 that does not suffer these problems, and `fgsea` a
fast implementation of the preranked gene set enrichment analysis (GSEA) algorithm
[@Korotkevich2016-gd].
For the ORA functions above, significantly differentially expressed genes with FDR 0.05 were
selected for the foreground gene set.
If fewer than 200 genes were detected as significant, then 200 genes with the smallest
p-values were used for ORA, like a previous report [@Tarca2013-xh].
For `fgsea`, the DESeq2 test statistic was used for scoring differential expression.
After enrichment analysis with these approaches, the significant up and down regulated
sets were selected and compared to the ground truth to calculate precision, recall
and F1 score.
At each value of added variance, 1000 replications were conducted.
No set seed was used, but results were similar when run on different computer systems.
All analyses were undertaken with R v4.4.0 in a Docker container.

## Results

### Understanding the background problem

A workaround to eliminate the background problem was developed and we used it to compare the
results of ORA of seven datasets with the original and corrected methods for nine different
gene set libraries.
The mean number of statistically significant results was calculated for each of the gene set
libraries before and after correcting the problem (**Figure 1A**).
Each of the gene set libraries showed an uplift when correcting the background problem.
The smallest increase was observed in Cellmarkers gene library (11.8%), while the largest
increase was observed for HPO (224%).
The similarity between results, as quantified with the Jaccard similarity, was highest for
Cellmarkers and GTRD at 0.76 and 0.77 respectively, and the lowest similarity was recorded for
HPO and KEGG with 0.20 and 0.48 respectively.
Observed Jaccard scores correlated with the number of genes annotated to one or more functional
categories in the gene set library (**Figure 1B**).
This result suggests that ORA with smaller gene set libraries like HPO and KEGG are impacted
more severely.

![**Figure 1A. Effect of correcting the background problem.** Figures represent the mean number
of statistically significant sets (FDR<0.05) over seven independent datasets detected with
clusterProfiler with and without correcting the background problem.
Separate tests for up- and down-regulated gene lists were conducted.
The percent increase in significant sets are shown.](../analysis/fig1_ORAsig.png "Figure 1A
shows that correcting the background bug improves the number of significant sets between 12 and
229%.")

![**Figure 1B. Impact of background problem is worse for smaller gene set libraries.** Mean
Jaccard similarity index was calculated for original and corrected ORA for seven independent
datasets. The number of genes represented in one or more functional sets is shown on the x-axis,
and the mean Jaccard index is shown on the y-axis.](../analysis/fig1_ORAjac.png "Figure 1B
shows that the background bug impacts smaller gene sets libraries like KEGG and HPO more
severely.")

### Understanding the FDR problem

A workaround for the FDR problem was developed, then original and corrected
analyses were conducted for the seven datasets and nine gene set libraries.
Results showed no major differences in the number of significant sets between
original and corrected analyses (**Figure 2A**).

We posited that the foreground gene list length might exacerbate the problem, so we performed
parallel analysis with foreground lists of length between 125 and 2,000 genes (**Figure 2B**).
A clear trend was observed for GO, Reactome, Cellmarkers and TFT GTRD, with Jaccard similarity
being significantly reduced when foreground gene lists were shorter than 500 genes.
Interestingly, the drop in Jaccard similarity was more severe for GO and Reactome sets as
compared to Cellmarkers and TFT GTRD sets and may be associated with the size of the gene
sets in the library.
Median gene set size for these are GO:18, Reactome:11, TFT GTRD:287 and Cellmarkers:115.
These results demonstrate that the FDR problem is more severe when dealing with shorter
foreground lists, and this is exacerbated when functional categories in the gene set library
are also small.

![**Figure 2A. Effect of correcting the FDR problem.** Figures represent the mean number of
significant sets with and without correction of the FDR problem over seven independent RNA-seq
datasets. Foreground list consisted of 2000 genes with the smallest p-values](../analysis/fig2_ORAfdr.png "Figure 2A shows that correcting the FDR bug slightly reduces the number of significant sets.")

![**Figure 2B. Impact of FDR problem is worse for shorter foreground gene lists.** Mean Jaccard
index between original and corrected analysis were calculated for seven RNA-seq datasets and nine
gene set libraries. Only GO, Reactome, Cellmarkers and TFT GTRD are shown.](
../analysis/fig2_ORAsz.png "Figure 2B shows that the impact of the FDR bug is worse when
foreground gene lists are shorter than 500.")

### Impact of these two problems on accuracy determined with simulated data

Simulated differential expression profiles with *a priori* changes to predetermined pathway
genes were analysed in parallel by clusterProfiler default and with mitigations for
problems described above.
In addition, these data underwent ORA with fora which does not suffer from these two
problems.
For comparison, fgsea was used to compare the performance of these two ORA methods against a
prototypical FCS method.
These methods were tested with different levels of introduced variation and the mean
precision, recall and F1 scores were recorded (**Figure 3A**).
Mitigating the background problem improved recall (by 1.8%), however it caused
a similar-sized decrease in precision (1.6% lower).
Addressing the FDR problem improved precision by 9.9%, but the recall dropped by 7.7%.
Addressing both problems improved precision by 8.2% but decreased recall by 5.4%.
Under these conditions, the FDR problem affected results more severely than the background
problem.
Results obtained for clusterProfiler after implementing the two mitigations were identical to
fora to three significant figures.
fgsea recorded precision of 0.940, similar to the ORA methods, but recall was far superior,
with fgsea scoring 22.9% higher than default clusterProfiler and 28.3% higher than fora.
As a result, fgsea's overall accuracy as defined by F1 index was 21.0% higher than fora and
18.6% higher than default clusterProfiler.

When looking at the accuracy profile at different levels of added noise (**Figure 3B**),
ORA techniques show high precision when the added noise is low (<0.2), but
precision drops at higher levels of added noise.
The rate of drop in precision was more severe for approaches that suffer from the FDR problem.
The precision of fgsea was relatively stable despite high levels of added noise, ranging from
0.95 to 0.92.
The recall of fgsea was superior to all ORA approaches across all levels of added noise.

![**Figure 3A. Impact of correcting these two problems on accuracy of simulated expression
data.**
Precision, recall and F1 scores are the mean of trials at a range of added noise.
CP; clusterProfiler, CP BG fix; clusterProfiler with a workaround to the background problem,
CP FDR fix; clusterProfiler with a workaround to the FDR problem,
CD BG and FDR fix; clusterProfiler with mitigations for both background and FDR problems,
fora; ORA function of the fgsea package,
fgsea; an FCS method similar to GSEA.
](../analysis/sim/fig3_bars.png "Figure 3A shows that fora is more
precise than clusterProfiler, and fgsea has higher recall than ORA methods.")

![**Figure 3B. Accuracy of clusterProfiler, fora and fgsea with simulated data with various
levels of added noise.**
Precision, recall and F1 score are shown with the addition of different amounts of random
noise to expression levels.
The purple line for clusterProfiler with mitigations for both problems is not visible as the
performance was identical to fora.
](../analysis/sim/fig3_sim.png "Figure 3B shows that ORA methods have deteriorating accuracy
with noisy data as compared to fgsea, a FCS method.")

## Discussion

The background problem came to our attention when a doctoral student in our department
was puzzled by results obtained from clusterProfiler when using a custom set of pathways.
This exact issue was the subject of several posts to online bioinformatics forums dating to
~2020 where confused users were wondering why the number of background genes changed so much.
The results shown here show that the background problem affects smaller gene set libraries
like KEGG and HPO more severely as compared to Cellmarkers and GO libraries which describe
a larger proportion of all genes.
ClusterProfiler maintainers recently provided an option to prevent the loss of unannotated
genes from the background `options(enrichment_force_universe = TRUE)`, however this feature
isn't yet described in the official documentation.

The second problem is potentially more serious in that it could lead to increasing false
positives under specific conditions; when the foreground list is short, and the median pathway
size is relatively small.
These false positives arise due to an underestimate of the actual number of tests conducted.
In extreme cases, correction of this problem can lead to around half of the statistically
significant results changing.

ClusterProfiler is not the only tool to suffer from these problems, they are widespread
(**Table 1**), but there are also several tools free of these problems.

| Tool | Has problems | Citation |
| --- | --- | --- |
| ClusterProfiler | Background and FDR | [@Yu2012-jj;@Wu2021-wy] |
| DAVID | Background and FDR | [@Dennis2003-ar;@Huang2009-gi;@Sherman2022-jn] |
| fora (fgsea) | Neither | [@Korotkevich2016-gd] |

: Table 1. Classification of several popular ORA tools for Background and FDR problems.

The purpose of this work was to describe two subtle problems with ORA, but there are some
bigger problems noted.
Firstly, ORA was less accurate as compared to FCS in this simulation work, confirming a
previous report [@Kaspi2020-rn].
Specifically, precision of ORA deteriorated with greater levels of added noise, while FCS
was relatively stable.
This means researchers can be more confident with FCS results, even if the data are
noisy.
Higher recall of FCS also means that researchers will have more power to identify pathways
relevant to the biological processes being studied.

The lack of enrichment scores hampers interpretation.

Lastly, ORA results can vary dramatically depending on the number of genes selected in the
foreground, a mostly arbitrary choice.
These issues together with previously mentioned pitfalls such as lack of background
correction and methodological misreporting [@Timmons2015-ki;@Wijesooriya2022-fw], suggest
that ORA should be discouraged in most situations, and replaced by other methods such as FCS.
Where it makes sense to use ORA would be when FCS is not applicable, for example when
investigating the over-represented categories in gene expression modules defined by packages
like WGCNA (Weighted Gene Correlation Network Analysis) [@Langfelder2008-hu], or when
investigating over-represented functional categories among genes with variants associated
with a disease.

# Data and Software Availability

## Underlying data

Publicly available data were obtained from Digital Expression Explorer 2 (dee2.io).

## Software and Code

* The code repository including template Dockerfile and R Markdown script are available on
GitHub (https://github.com/markziemann/background).

* The example Docker image is available on DockerHub
(https://hub.docker.com/r/mziemann/background).

* The code repository and Docker image have been uploaded to Zenodo for long-term archiving
(https://zenodo.org/record/8170984).

## Availability of other materials

## Author Contributions

Conceptualization: MZ, AB.
Data Curation: MZ.
Formal Analysis: MZ.
Funding Acquisition: N/A.
Investigation: MZ, AB.
Methodology: MZ, AB.
Project Administration: MZ.
Resources: MZ.
Software: MZ, AB.
Supervision: MZ.
Validation: AB.
Visualization: MZ, AB.
Writing – Original Draft Preparation: MZ, AB.
Writing – Review & Editing: MZ, AB.

## Competing Interests

No competing interests were disclosed.

## Grant Information

The authors declared that no grants were involved in supporting this work.

## Acknowledgements

We thank Ms Megan Soria and Ms Kaumadi Wijesooriya (Deakin University) for discussions on the concept and manuscript.
This research was supported by use of the Nectar Research Cloud, a collaborative Australian research
platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).
The authors gratefully acknowledge the contribution to this work of the
Victorian Operational Infrastructure Support Program received by the Burnet Institute.

# Bibliography