### 2024-07-01 app/report.Rmd
FDR fix

---
title: "Two subtle issues with over-representation analysis: Demonstration"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
params:
  fg: NA
  bg: NA
  comparison: NA
  genesetlibrary: NA
  mygs: NA
---

This software was developed by Mark Ziemann under an MIT licence.
Report issues at the [GitHub repository](https://github.com/markziemann/background).

## Background

Over-representation analysis is frequently used to determine the functionally
important themes from gene lists.
These gene lists are classified in different ways, but are commonly defined
by differential regulation in omics experiments like RNA-seq, proteomics or
micro-array.
Many different tools are available to conduct this type of analysis including
[DAVID](https://david.ncifcrf.gov/) and [Enricher](https://maayanlab.cloud/Enrichr/).

For the results to be reliable, the gene list needs to be provided together
with a background list.
The background list is the full list of genes that were reliably detected by
the assay.
This is important, because these assays typically cannot detect all genes
reliably due to technical limitations.
As different tissues express genes very differently, the right background list
is critical to obtain the functional categories that are specific to the
experiment conducted.
Using a generic background list of all annotated genes could lead to a major
bias.

Additionally, as such analyses involve the analysis of hundreds of
functional categories in parallel, the resulting p-values cannot be taken at
face value due to the high rate of false positives.
To mitigate this, false discovery rate correction is applied to adjust the
p-values.

In this work we highlight two subtle issues with some popular implementations 
of over-representation analysis.

### Background problem: Genes without annotated categories are discarded

The first issue related to how the background list is defined.
If a user does the right thing and provides a background list along with the
foreground list, then some tools will exclude any genes from the background
that do not have any annotated functional categories.
This results in a large fraction of genes in the background being excluded,
and the enrichment ratio being unreliable.
This issue is more severe for analyses that involve smaller libraries of
functional annotations.
For example KEGG has annotations for a few thousand genes, meaning most of
the background genes get discarded.

### FDR problem: p-value adjustment is applied incorrectly

The second issue is that the false discovery rate is implemented incorrectly.
ORA enrichment tools intersect the gene list with each functional category
and return results for each functional category with at least one common
gene.
But there are some functional categories with zero genes in common with the
gene list.
These are excluded from the end results.
The process of attempting to intersect these categories with the submitted
gene list is itself a test, and so statistically speaking, the result of
that failed test should be reported in the final results with no overlaps and
a p-value of 1.
This may not be a problem when running enrichment analysis on a gene list 
with 3000 members, but it might be serious when it is small, like <200
members.
This results in adjusted p-values being slightly smaller than they should be.

```{r, start}

fg <- params$fg
bg <- params$bg
comparison <- params$comparison
genesetlibrary <- params$genesetlibrary
mygs <- params$mygs
minsetsize = 5

suppressPackageStartupMessages({
  library("clusterProfiler")
  library("markdown")
  library("rmarkdown")
  library("shiny")
  library("broom")
  library("RhpcBLASctl")
  library("eulerr")
  library("plotly")
  library("DT")
})

```

### Notes

Minimum set size is defined as 5.
Sets with fewer than this number of genes are discarded.

The problem we are characterising is:

```{r,notes}

mymessage <- paste("Here we are characterising",comparison)
message(mymessage)

```

## Input gene lists

First 10 genes in the foreground.

```{r,peek1}
head(fg,10)
```

First 10 genes in the background.

```{r,peek2}
head(bg,10)
```

Number of genes in the foreground list.

```{r,peek3}
length(fg)
```

Number of genes in the background list.

```{r,peek4}
length(bg)
```

## Gene set library

Here is a look at the gene set library.

```{r, gsl1}
genesetlibrary
```

Number of sets.

```{r,gsl1a}
length(unique(mygs$term))
```

Summary of gene set length.

```{r,gsl2}
summary(as.numeric(unname(table(mygs$term))))
```

Total number of genes with one or more annotations in the library.

```{r,gsl3}
length(unique(mygs$gene))
```

Total number of annotations in the gene set library.

```{r,gsl4}
nrow(mygs)
```

Overlap of foreground genes in the gene set library.

```{r,gsl5}

mygs_genes <- unique(mygs$gene)

I <- length(intersect(fg,mygs_genes))

message(paste(I,"from",length(fg)))

```

Overlap of background genes in the gene set library.

```{r,gsl6}

I <- length(intersect(bg,mygs_genes))

message(paste(I,"from",length(bg)))

```

## Default ORA

```{r,default }

options(enrichment_force_universe = FALSE)
ora <- as.data.frame(enricher(gene = fg ,
                              universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = mygs,
                              pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

ora$ES <- gr/br
ora$ID = ora$geneID = ora$p.adjust = ora$Count = NULL
colnames(ora) <- gsub("qvalue","FDR",colnames(ora))
colnames(ora) <- gsub("GeneRatio","FgRatio",colnames(ora))
ora <- ora[,c("Description","FgRatio","BgRatio","ES","pvalue","FDR")]
ora$ES <- signif(ora$ES,3)
ora$pvalue <- signif(ora$pvalue,3)
ora$FDR <- signif(ora$FDR,3)
ora1 <- ora

```

```{r,defaulttable}

DT::datatable(
  { ora1 },

  extensions = 'Buttons',
  filter = list(position = 'top', clear = FALSE),

  options = list(
    paging = TRUE,
    searching = TRUE,
    fixedColumns = TRUE,
    autoWidth = TRUE,
    ordering = TRUE,
    dom = 'tB',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 50,
    search = list(regex = TRUE, caseInsensitive = TRUE)),

  rownames= FALSE)

```

Bar chart.

```{r,default_bars}

edata <- head(subset(ora1,FDR<0.05),20)
if (nrow(edata)>1) {
  edata <- edata[order(edata$ES),]
  vec <- edata$ES
  names(vec) <- edata$Description
  names(vec) <- substr(names(vec), start = 1, stop = 60)
  names(vec) <- gsub("_"," ",names(vec))
  par(mar=c(4,27,1,1))
  barplot(vec,horiz=TRUE,las=1,cex.names=0.8,
          xlab="Enrichment Score")
  mtext("Default ORA",side=3,adj=1)
} else {
  message("No gene sets with FDR<0.05 found")
}

```

## Fixed analysis

There are two issues with ORA.
The one being characterised here is shown in the text below:

```{r,fixed1 }

mymessage <- paste("Here we are characterising",comparison)
message(mymessage)

#Background error

if (comparison == "Background error") {
  options(enrichment_force_universe = TRUE)
  ora <- as.data.frame(enricher(gene = fg ,
                                universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = mygs,
                                pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

  gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

  ora$ES <- gr/br
  ora$ID = ora$geneID = ora$p.adjust = ora$Count = NULL
  colnames(ora) <- gsub("qvalue","FDR",colnames(ora))
  colnames(ora) <- gsub("GeneRatio","FgRatio",colnames(ora))
  ora <- ora[,c("Description","FgRatio","BgRatio","ES","pvalue","FDR")]
  ora$ES <- signif(ora$ES,3)
  ora$pvalue <- signif(ora$pvalue,3)
  ora$FDR <- signif(ora$FDR,3)
} else if (comparison == "FDR error") {
  gs1 <- mygs
  gs1 <- gs1[which(gs1$gene %in% bg),]
  terms <- names(which(table(gs1$term)>5))
  gs1 <- gs1[gs1$term %in% terms,]
  options(enrichment_force_universe = FALSE)
  ora <- as.data.frame(enricher(gene = fg ,
                                universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = gs1,
                                pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

  gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

  ora$ES <- gr/br
  ora$ID = ora$geneID = ora$p.adjust = ora$Count = NULL
  colnames(ora) <- gsub("qvalue","FDR",colnames(ora))
  colnames(ora) <- gsub("GeneRatio","FgRatio",colnames(ora))
  ora <- ora[,c("Description","FgRatio","BgRatio","ES","pvalue","FDR")]

  # proper FDR correction
  nsets <- length(which(table(gs1$term)>5))
  nres <- nrow(ora)
  diff <- nsets - nres
  pvals <- c(ora$pvalue,rep(1,diff))
  ora$FDR <- p.adjust(pvals,method="fdr")[1:nrow(ora)]
}

ora2 <- ora
```

```{r,fixed2}

DT::datatable(
  { ora2 },

  extensions = 'Buttons',
  filter = list(position = 'top', clear = FALSE),

  options = list(
    paging = TRUE,
    searching = TRUE,
    fixedColumns = TRUE,
    autoWidth = TRUE,
    ordering = TRUE,
    dom = 'tB',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 50,
    search = list(regex = TRUE, caseInsensitive = TRUE)),

  rownames= FALSE)

```

Fixed bar chart.

```{r,fixed_bars}

edata <- head(subset(ora2,FDR<0.05),20)
if (nrow(edata)>1) {
  edata <- edata[order(edata$ES),]
  vec <- edata$ES
  names(vec) <- edata$Description
  names(vec) <- substr(names(vec), start = 1, stop = 60)
  names(vec) <- gsub("_"," ",names(vec))
  par(mar=c(4,27,1,1))
  barplot(vec,horiz=TRUE,las=1,cex.names=0.8,
          xlab="Enrichment Score")
  mtext("Fixed ORA",side=3,adj=1)
} else {
  message("No gene sets with FDR<0.05 found")
}

```

## Comparison of default and fixed analysis

Side by side data table.
Columns labeled 'x' are original algorithm, and 'y' are fixed.

```{r,comparison1}
m <- merge(ora1,ora2,by="Description")
m <- m[,c("Description","FgRatio.x","FgRatio.y","BgRatio.x","BgRatio.y","ES.x","ES.y","FDR.x","FDR.y")]
diff <- abs(-log10(m$FDR.x) - -log10(m$FDR.y))
m <- m[order(-diff),]
m$ES.x <- signif(m$ES.x,3)
m$ES.y <- signif(m$ES.y,3)
m$FDR.x <- signif(m$FDR.x,3)
m$FDR.y <- signif(m$FDR.y,3)

DT::datatable( { m },
  extensions = 'Buttons',
  filter = list(position = 'top', clear = FALSE),
  options = list(
    paging = TRUE,
    searching = TRUE,
    fixedColumns = TRUE,
    autoWidth = TRUE,
    ordering = TRUE,
    dom = 'tB',
    buttons = c('copy', 'csv', 'excel'),
    pageLength = 50,
    search = list(regex = TRUE, caseInsensitive = TRUE)),
  rownames= FALSE)

```

Euler diagram of pathways with FDR<0.05.

```{r,comparison2}

ora1_sets <- subset(ora1,FDR < 0.05)$Description
ora2_sets <- subset(ora2,FDR < 0.05)$Description

if ( comparison == "Background error") {
  v1 <- list("Original"=ora1_sets, "BG fix"=ora2_sets)
} else if (comparison == "FDR error") {
  v1 <- list("Original"=ora1_sets, "FDR fix"=ora2_sets)
}
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2))

```

Jaccard similarity index of pathways with FDR<0.05.

```{r,jac1}

jac <- length(intersect(ora1_sets,ora2_sets)) / length(union(ora1_sets,ora2_sets))

message(paste("Jaccard index:",signif(jac,3)))

```

Scatterplot of ES.

```{r,comparison3}

orig_df <- ora1
bgfix_df <- ora2
m <- merge(orig_df,bgfix_df,by="Description")
m <- m[,c("Description","ES.x","ES.y")]
MAX <- max(c(m$ES.x,m$ES.y))
INCREMENT <- MAX/nrow(m)
VEC <- seq(0,MAX,INCREMENT)

if ( comparison == "Background error") {

  m$Original <- m$BGcorrected <- round(VEC[1:nrow(m)],0.1)
  m$ES.x <- signif(m$ES.x,3)
  m$ES.y <- signif(m$ES.y,3)

  plot_ly(
    m, x = ~ES.x, y = ~ES.y
  ) %>%
    add_trace(m, x = ~Original, y = ~BGcorrected, type = "scatter",
              mode="lines", line=list(color='grey')) %>%
    add_trace(m, x = ~ES.x, y = ~ES.y, type = "scatter", mode = "markers",
              showlegend = FALSE, text = m$Description,
              hoverinfo = 'text') %>%
    layout(title="Enrichment score comparison")
} else if (comparison == "FDR error") {
  m$Original <- m$FDRcorrected <- round(VEC[1:nrow(m)],0.1)
  m$ES.x <- signif(m$ES.x,3)
  m$ES.y <- signif(m$ES.y,3)

  plot_ly(
    m, x = ~ES.x, y = ~ES.y
  ) %>%
    add_trace(m, x = ~Original, y = ~FDRcorrected, type = "scatter",
              mode="lines", line=list(color='grey')) %>%
    add_trace(m, x = ~ES.x, y = ~ES.y, type = "scatter", mode = "markers",
              showlegend = FALSE, text = m$Description,
              hoverinfo = 'text') %>%
    layout(title="Enrichment score comparison")
}

```

Scatterplot of logFDR values.

```{r,comparison4}

m <- merge(ora1,ora2,by="Description")
m <- m[,c("Description","FDR.x","FDR.y")]
m$logFDR.x <- -log10(m$FDR.x)
m$logFDR.y <- -log10(m$FDR.y)
MAX <- max(c(m$logFDR.x,m$logFDR.y))
INCREMENT <- MAX/nrow(m)
VEC <- seq(0,MAX,INCREMENT)

if ( comparison == "Background error") {

  m$Original <- m$BGcorrected <- round(VEC[1:nrow(m)],0.1)
  m$logFDR.x <- signif(m$logFDR.x,3)
  m$logFDR.y <- signif(m$logFDR.y,3)

  plot_ly(
    m, x = ~logFDR.x, y = ~logFDR.y
  ) %>%
    add_trace(m, x = ~Original, y = ~BGcorrected, type = "scatter",
              mode="lines", line=list(color='grey')) %>%
    add_trace(m, x = ~logFDR.x, y = ~logFDR.y, type = "scatter", mode = "markers",
              showlegend = FALSE, text = m$Description,
              hoverinfo = 'text') %>%
    layout(title="log FDR comparison")

} else if (comparison == "FDR error") {
  m$Original <- m$FDRcorrected <- round(VEC[1:nrow(m)],0.1)
  m$logFDR.x <- signif(m$logFDR.x,3)
  m$logFDR.y <- signif(m$logFDR.y,3)

  plot_ly(
    m, x = ~logFDR.x, y = ~logFDR.y
  ) %>%
    add_trace(m, x = ~Original, y = ~FDRcorrected, type = "scatter",
              mode="lines", line=list(color='grey')) %>%
    add_trace(m, x = ~logFDR.x, y = ~logFDR.y, type = "scatter", mode = "markers",
              showlegend = FALSE, text = m$Description,
              hoverinfo = 'text') %>%
    layout(title="log FDR comparison")
}

```

## Session information

Here is the session info with all the versions of packages used.

```{r session_info}

sessionInfo()

```

END of report