### 2024-07-17 analysis/GSE158420_rna.Rmd
plotting

---
title: "GSE158420: RNA expression downsampling"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/background/

## Introduction

Here, we look at the sensitivity of pathway analysis of RNA-seq data.

```{r,packages}

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("HGNChelper")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("gplots")
  library("gridExtra")
  library("png")
  library("fgsea")
  library("parallel")
})

CORES=16

```

## Load annotations and methylation data

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs <- gmt_import("../gmt/c2.cp.reactome.v2023.2.Hs.symbols.gmt")

names(gs) <- gsub("REACTOME_","",names(gs))
names(gs) <- gsub("_"," ",names(gs))

```

## Load RNA expression

From GEO https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158420
GSE158420_counts.txt.gz

```{r,deseq1}

if (!file.exists("GSE158420_counts.txt.gz")) {
  download.file("https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE158420&format=file&file=GSE158420%5Fcounts%2Etxt%2Egz",
    destfile="GSE158420_counts.txt.gz")
}

x <- read.table("GSE158420_counts.txt.gz",row.names=NULL)
dim(x)
x <- aggregate(. ~ row.names,x,sum)
dim(x)

#RUFKM
head(x)

```

There are some gene name errors, so I will use HGNChelper to repair them.

```{r,gene_name_fix}

human=x$row.names
new.hgnc.table <- readRDS("new.hgnc.table.rds")
fix <- checkGeneSymbols(human,map=new.hgnc.table)
x$gene <- fix$Suggested.Symbol
x$row.names=NULL
x <- aggregate(. ~ gene,x,sum)
dim(x)
row.names(x)=x$gene
x$gene=NULL
head(x)

```

Now do DESeq2

```{r,deseq2}

patient <- head(as.character(unlist(lapply(1:ncol(x),function(i) {c(i,i)}))),ncol(x))
tumor <- rep(c(0,1),ncol(x)/2)
ss <- data.frame(patient,tumor)
rownames(ss) <- colnames(x)

dds <- DESeqDataSetFromMatrix(countData = x , colData=ss, design = ~ patient + tumor)
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),assay(vsd))
dge <- as.data.frame(zz[order(zz$pvalue),])

sig <- subset(dge,padj<0.05)
SIG = nrow(sig)
DN = nrow(subset(sig,log2FoldChange<0))
UP = nrow(subset(sig,log2FoldChange>0))
HEADER = paste("mRNA", SIG , "DEGs,", UP ,"upregulated,", DN, "downregulated")
# smear
plot(log2(dge$baseMean),dge$log2FoldChange,cex=0.6,cex.axis=1.2,cex.lab=1.3,
  xlab="log2 base mean",
  ylab="log2 fold change" ,pch=19,col="#838383")
points(log2(sig$baseMean),sig$log2FoldChange,cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
top <- head(sig,20)
# volcano
plot(dge$log2FoldChange, -log2(dge$pvalue) ,cex=0.6, cex.lab=1.3,cex.axis=1.2,
 ,xlab="log2 fold change", ylab="-log2 p-value" ,pch=19,col="#838383")
points(sig$log2FoldChange, -log2(sig$pvalue),cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
# top N gene heatmap
colCols <- rep(c(0,1),ncol(x)/2)
colCols <- gsub("0","gray",colCols)
colCols <- gsub("1","black",colCols)

colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(  as.matrix(dge[1:50,c(7:ncol(dge))]), col=colfunc(25),scale="row",
  ColSideColors=colCols ,
  trace="none",margins = c(6,10), cexRow=.6, cexCol=.5, main="Top 50 genes")

```

RNA expression pathway enrichment with mitch

```{r,mitch}

rna <- mitch_import(dge, DEtype="deseq2")

res <- mitch_calc(rna, genesets=gs, priority="effect",cores=16, minsetsize=5)

mres <- res$enrichment_result
msig <- subset(mres,p.adjustANOVA<0.05)
nrow(msig)

head(subset(mres,p.adjustANOVA<0.05 & s.dist>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(mres,p.adjustANOVA<0.05 & s.dist<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

```

FGSEA

```{r,fgsea}

dge$stat[which(is.na(dge$stat))] <- 1

stat <- dge$stat
names(stat) <- rownames(dge)

pll <- lapply(1:32,function(i) {
  tic()
  fres <- fgsea(pathways=gs,stats=stat,minSize=5, nproc=i)
  tm <- toc()
  tm$callback_msg
})

pll <- unlist(pll)
pll <- as.numeric(sapply(strsplit(pll," "),"[[",1))

names(pll) <- 1:32
barplot(pll)
plot(1:32,pll,type="b",pch=19)

fsig <- subset(fres,padj<0.05)
fres <- fres[order(-abs(fres$ES)),]

head(subset(fres,padj<0.05 & ES>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(fres,padj<0.05 & ES<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

fsets <- paste(fsig$pathway,as.character(fsig$ES > 0))

```

FORA

```{r,fora1}

sigup <- rownames(subset(dge,padj<0.05 & log2FoldChange>0))
sigdn <- rownames(subset(dge,padj<0.05 & log2FoldChange<0))
bg <- rownames(dge)

oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
oresup$es <- (oresup$overlap / length(sigup) ) / (oresup$size / length(bg))
head(oresup,10)

oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
oresdn$es <- (oresdn$overlap / length(sigdn) ) / (oresdn$size / length(bg))
head(oresup,10)

oresup <- subset(oresup,padj<0.05)
oresup <- paste(oresup$pathway,"TRUE")

oresdn <- subset(oresdn,padj<0.05)
oresdn <- paste(oresdn$pathway,"FALSE")

osets <- c(oresup,oresdn)
str(osets)

```

Assess the effect of gene number on pathway results.

```{r,fora2}

sizes <- c(100,200,500,1000,2000,3000,4000,5000,6000,7000,8000)

ores <- lapply(sizes, function(size) {
  sigup <- head(rownames(subset(dge,padj<0.05 & log2FoldChange>0)),size)
  sigdn <- head(rownames(subset(dge,padj<0.05 & log2FoldChange<0)),size)
  bg <- rownames(dge)

  oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
  oresup$es <- (oresup$overlap / length(sigup) ) / (oresup$size / length(bg))
  nup <- nrow(subset(oresup,padj<0.05))

  oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
  oresdn$es <- (oresdn$overlap / length(sigdn) ) / (oresdn$size / length(bg))
  ndn <- nrow(subset(oresdn,padj<0.05))

  return(c(nup,ndn))
})

ores <- do.call(rbind,ores)
rownames(ores) <- sizes
colnames(ores) <- c("up","dn")
ores <- as.data.frame(ores)
ores$tot <- ores$up + ores$dn
tot <- fres$tot
names(tot) <- sizes
par(mar=c(mar = c(5.1, 5.1, 4.1, 2.1)))
barplot(tot,horiz=TRUE,las=1,xlab="no. pathways FDR<0.05",ylab="no. genes selected")

```

## Downsample FGSEA

This will run the LAM analysis at group sizes of 2, 3, 6, 10, 16, 22 and 30,
50 times each, with pseudorandom seeds from 100 to 5000 jumping up by 100.

```{r,fgsea_run,fig.height=10}

RhpcBLASctl::blas_set_num_threads(1) #set automatic cores to 1

SAMPLESIZES=c(2,3,5,10,15,20,25,30)
#SAMPLESIZES=c(3,6,10,22)
fgres <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100) # 50 repetitions
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    ss2 <- ss[mysample,]
    x2 <- x[,mysample]
    x2 <- x2[rowMeans(x2)>=10,]

    dds <- DESeqDataSetFromMatrix(countData = x2 , colData=ss2, design = ~ patient + tumor)
    res <- DESeq(dds)
    z <- results(res)
    vsd <- vst(dds, blind=FALSE)
    zz <-cbind(as.data.frame(z),assay(vsd))
    dge <- as.data.frame(zz[order(zz$pvalue),])

    dge$stat[which(is.na(dge$stat))] <- 1
    stat <- dge$stat
    names(stat) <- rownames(dge)
    fres <- fgsea(pathways=gs,stats=stat,minSize=5,nproc=3)
    fres <- fres[order(-abs(fres$ES)),]
    fsig <- subset(fres,padj<0.05)
    fsets2 <- paste(fsig$pathway,as.character(fsig$ES > 0))

    TOT=length(fsets2)
    TP=length(which(fsets2 %in% fsets))
    FP=length(which(!fsets2 %in% fsets))
    FN=length(which(!fsets %in% fsets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=16)
  do.call(rbind,dsres)
})

ssres <- fgres
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

pdf("sens_fgsea.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

dev.off()

ssres <- fgres
dots <- lapply(ssres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

```

## FORA Downsample

```{r,fora_run,fig.height=10}

SAMPLESIZES=c(2,3,5,10,15,20,25,30)
fores <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100) # 50 repetitions
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    ss2 <- ss[mysample,]
    x2 <- x[,mysample]
    x2 <- x2[rowMeans(x2)>=10,]

    dds <- DESeqDataSetFromMatrix(countData = x2 , colData=ss2, design = ~ patient + tumor)
    res <- DESeq(dds)
    z <- results(res)
    vsd <- vst(dds, blind=FALSE)
    zz <-cbind(as.data.frame(z),assay(vsd))
    dge <- as.data.frame(zz[order(zz$pvalue),])

    sigup <- rownames(subset(dge,padj<0.05 & log2FoldChange>0))
    if (length(sigup) < 250) { sigup <- head(rownames(subset(dge, log2FoldChange>0)),250) }
    sigdn <- rownames(subset(dge,padj<0.05 & log2FoldChange<0))
    if (length(sigdn) < 250) { sigdn <- head(rownames(subset(dge, log2FoldChange<0)),250) }
    bg <- rownames(dge)

    oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
    oresup <- subset(oresup,padj<0.05)
    oresup <- paste(oresup$pathway,"TRUE")

    oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
    oresdn <- subset(oresdn,padj<0.05)
    oresdn <- paste(oresdn$pathway,"FALSE")

    osets2 <- c(oresup,oresdn)

    TOT=length(osets2)
    TP=length(which(osets2 %in% osets))
    FP=length(which(!osets2 %in% osets))
    FN=length(which(!osets %in% osets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=16)
  do.call(rbind,dsres)
})

ssres <- fores

par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

pdf("sens_fora.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

dev.off()

## Summarise downnsampling

```{r,downsample,fig.width=10}

par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
mtext("no. consistent pathways FDR<0.05")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
mtext("no. inconsistent pathways FDR<0.05")
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
mtext("Proportion of inconsistent pathways")
grid()

### PDF
pdf("sens.pdf",height=4,width=8)
par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", main="no. consistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",main="no. inconsistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",main="Proportion of inconsistent pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()
dev.off()

```
## Session information

```{r,save}

save.image("GSE158422_sensitivity.Rdata")

sessionInfo()

```