### 2024-09-09 analysis/analyse_rnaseq1.Rmd
adding error bars

---
title: "Examining background gene list issue in clusterProfiler - RNA-seq example"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/background

## Introduction

This guide is a Rmarkdown script that conducts differential expression and enrichment analysis,
which are very popular workflows for transcriptome data.

The goal of this work is to understand how ClusterProfiler manages the background list,
as we have observed some weird behaviour.
In particular we see that when we provide a custom background, those genes do not appear in the
final analysis unless they are also included in the gene list.

In the code chunk below called `libs`, you can add and remove required R library dependancies.
Check that the libraries listed here match the Dockerfile, otherwise you might get errors.

```{r,libs}

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
  library("parallel")
})

```

For this guide I will be using bulk RNA-seq data from several previous studies.

## Load data

Gene sets were obtained from MSigDB.

```{r,load_data}

myfiles <- list.files("../dataprep",pattern="*Rds",full.names=TRUE)
d <- lapply(myfiles,readRDS)
names(d) <- basename(myfiles)

mygmts <- list.files("../gmt",pattern="*gmt",full.names=TRUE)
gs <- lapply(mygmts,gmtPathways)
names(gs) <- basename(mygmts)

names(gs)
names(gs) <- c("KEGG", "Reactome", "Wikipathways", "miR targets", "TFT GTRD",
  "GO", "HPO", "Cellmarkers", "Hallmark" )

```

## Basic geneset analysis

```{r,gs1}

# number of sets
num <- unlist(lapply(gs,length))
num

# set size summary
sz <- unlist(lapply(gs,function(x) {
  y <- unlist(lapply(x, length ))
  summary(y)
} ))
sz <- t(matrix(sz,ncol=9))
rownames(sz) <- names(gs)
colnames(sz) <- names(summary(c(1,2,3,4)))
sz

# gene coverage
cov <- unlist(lapply(gs,function(x) {
  y <- unique(unlist(x))
  length(y)
} ))
cov

# total number of annotations
tot <- unlist(lapply(gs, function(x) {
  sum(unlist(lapply(x,length)))
} ))
tot

```

## First test

```{r,initialtest1}

x1 <- readRDS("../dataprep/bulkrna1.Rds")
head(x1)

# remove genes without symbols
x1f <- x1[lapply(strsplit(rownames(x1) , " "),length)>1,]

dim(x1)
dim(x1f)

# about 1000 genes are lost after removing redundant gene names
bg <- unique(sapply(strsplit(rownames(x1f) , " "),"[[",2))
length(bg)

up <-rownames(subset(x1f, log2FoldChange > 0 & padj < 0.05 ))
up <- unique(sapply(strsplit(up," "),"[[",2))
length(up)

dn <-rownames(subset(x1f, log2FoldChange < 0 & padj < 0.05 ))
dn <- unique(sapply(strsplit(dn," "),"[[",2))
length(dn)

# kegg gene set
kegg <- gs[[1]]
kegg <- stack(kegg)
colnames(kegg) <- c("gene","term")
kegg <- kegg[,c(2,1)]
kegg$term <- gsub("KEGG_MEDICUS_","",kegg$term)
kegg$term <- gsub("_"," ",kegg$term)

ora_up <- as.data.frame(enricher(gene = up ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = kegg,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

 gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

  ora_up$es <- gr/br

ora_up$geneID = ora_up$ID = ora_up$Description = NULL

ora_up_kegg <- ora_up

head(ora_up,20) %>%
  kbl(row.names = TRUE, caption="Upregulated KEGGs in SRP128998") %>%
  kable_paper("hover", full_width = F)

# GO gene set
go <- gs[[6]]
go <- stack(go)
colnames(go) <- c("gene","term")
go <- go[,c(2,1)]
go$term <- gsub("^GO","",go$term)
go$term <- gsub("_"," ",go$term)

ora_up <- as.data.frame(enricher(gene = up ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = go,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

 gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

  ora_up$es <- gr/br

ora_up$geneID = ora_up$ID = ora_up$Description = NULL

ora_up_go <- ora_up

head(ora_up,20) %>%
  kbl(row.names = TRUE, caption="Upregulated GOs in SRP128998") %>%
  kable_paper("hover", full_width = F)

```

Transfer a gene set from GO to KEGG analysis.

```{r,intitialtest2}

pw <- gs[[6]][grep("GOBP_COMPLEMENT_ACTIVATION_CLASSICAL_PATHWAY", names(gs[[6]]))]

pw

pw <- stack(pw)[,c(2,1)]
colnames(pw) <- c("term","gene")
kegg2 <- rbind(kegg,pw)

ora_up <- as.data.frame(enricher(gene = up ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = kegg2,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

 gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

  ora_up$es <- gr/br

ora_up$geneID = ora_up$ID = ora_up$Description = NULL

ora_up_kegg2 <- ora_up

head(ora_up,20) %>%
  kbl(row.names = TRUE, caption="Upregulated KEGGs in SRP128998 with Complement") %>%
  kable_paper("hover", full_width = F)

message("GO")
ora_up_go[rownames(ora_up_go)=="BP COMPLEMENT ACTIVATION CLASSICAL PATHWAY",]

message("KEGG + complement")
ora_up_kegg2[rownames(ora_up_kegg2)=="GOBP_COMPLEMENT_ACTIVATION_CLASSICAL_PATHWAY",]

```

Fix KEGG analysis.

```{r,keggfix1}

bgdf <- data.frame("background",bg)
colnames(bgdf) <- c("term","gene")
keggfix <- rbind(kegg,bgdf)

ora_up <- as.data.frame(enricher(gene = up ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = keggfix,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

 gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

  ora_up$es <- gr/br

ora_up$geneID = ora_up$ID = ora_up$Description = NULL

ora_up_keggfix <- ora_up

head(ora_up,20) %>%
  kbl(row.names = TRUE, caption="Upregulated KEGG fixed in SRP128998 with Complement") %>%
  kable_paper("hover", full_width = F)

message("Original")
ora_up_kegg[which(rownames(ora_up_kegg)=="REFERENCE CLASSICAL PATHWAY OF COMPLEMENT CASCADE C4 C2 TO C3 CONVERTASE FORMATION"),]

message("Fixed")
ora_up_keggfix[which(rownames(ora_up_keggfix)=="REFERENCE CLASSICAL PATHWAY OF COMPLEMENT CASCADE C4 C2 TO C3 CONVERTASE FORMATION"),]

```

## Enrichment with Clusterprofiler

Here is a function that conducts over-representation analysis with ClusterProfiler.
Genes that are not annotated as having any gene set category are discarded.

```{r,orabad}

# df = deseq output
# df must have "ensID geneSymbol" as rownames

# gs = gene sets
# gs format is list of symbols
orabad <- function( df, gs ) {

  def <- df

  gs1 <- stack(gs)
  colnames(gs1) <- c("gene","term")
  gs1 <- gs1[,c(2,1)]

  defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
  defup <- defup[lapply(strsplit(defup," "),length)==2]
  defup <- unique(sapply(strsplit(defup," "),"[[",2))

  if ( length(defup)<200 ) {
    defup <- head(rownames(subset(def,log2FoldChange>0)),200)
    defup <- defup[lapply(strsplit(defup," "),length)==2]
    defup <- unique(sapply(strsplit(defup," "),"[[",2))
  }

  defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
  defdn <- defdn[lapply(strsplit(defdn," "),length)==2]
  defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

  if ( length(defdn)<200 ) {
    defdn <- head(rownames(subset(def,log2FoldChange<0)),200)
    defdn <- defdn[lapply(strsplit(defdn," "),length)==2]
    defdn <- unique(sapply(strsplit(defdn," "),"[[",2))
  }

  bg <- rownames(def)
  bg <- bg[lapply(strsplit(bg," "),length)==2]
  bg <- unique(sapply(strsplit(bg," "),"[[",2))

  ora_up <- as.data.frame(enricher(gene = defup ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = gs1,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

  ora_up$geneID <- NULL
  ora_ups <- rownames(ora_up)

  gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

  ora_up$es <- gr/br
  ora_up <- ora_up[order(-ora_up$es),]
  ora_up$Description=NULL

  ora_dn <- as.data.frame(enricher(gene = defdn ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = gs1,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

  ora_dn$geneID <- NULL
  ora_dns <- rownames(ora_dn)

  gr <- as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",2))

  ora_dn$es <- gr/br
  ora_dn <- ora_dn[order(-ora_dn$es),]
  ora_dn$Description=NULL

  result <- list("orabad_up"=ora_up,"orabad_dn"=ora_dn)
  return(result)
}

```

This function fixes the problem with excluding un-annotated genes.

```{r,oragood}

oragood <- function( df, gs ) {

  def <- df

  gs1 <- stack(gs)
  colnames(gs1) <- c("gene","term")
  gs1 <- gs1[,c(2,1)]

  defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
  defup <- defup[lapply(strsplit(defup," "),length)==2]
  defup <- unique(sapply(strsplit(defup," "),"[[",2))

  if ( length(defup)<200 ) {
    defup <- head(rownames(subset(def,log2FoldChange>0)),200)
    defup <- defup[lapply(strsplit(defup," "),length)==2]
    defup <- unique(sapply(strsplit(defup," "),"[[",2))
  }

  defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
  defdn <- defdn[lapply(strsplit(defdn," "),length)==2]
  defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

  if ( length(defdn)<200 ) {
    defdn <- head(rownames(subset(def,log2FoldChange<0)),200)
    defdn <- defdn[lapply(strsplit(defdn," "),length)==2]
    defdn <- unique(sapply(strsplit(defdn," "),"[[",2))
  }

  bg <- rownames(def)
  bg <- bg[lapply(strsplit(bg," "),length)==2]
  bg <- unique(sapply(strsplit(bg," "),"[[",2))

  bgdf <- data.frame("background",bg)
  names(bgdf) <- c("term","gene")
  gs1 <- rbind(gs1,bgdf)

  ora_up <- as.data.frame(enricher(gene = defup ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = gs1,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

  ora_up$geneID <- NULL
  ora_ups <- rownames(ora_up)

  gr <- as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_up$BgRatio,"/"),"[[",2))

  ora_up$es <- gr/br
  ora_up <- ora_up[order(-ora_up$es),]
  ora_up$Description=NULL

  ora_dn <- as.data.frame(enricher(gene = defdn ,
    universe = bg, minGSSize = 5, maxGSSize = 500000, TERM2GENE = gs1,
    pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

  ora_dn$geneID <- NULL
  ora_dns <- rownames(ora_dn)

  gr <- as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_dn$GeneRatio,"/"),"[[",2))

  br <- as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",1)) /
    as.numeric(sapply(strsplit(ora_dn$BgRatio,"/"),"[[",2))

  ora_dn$es <- gr/br
  ora_dn <- ora_dn[order(-ora_dn$es),]
  ora_dn$Description=NULL

  result <- list("oragood_up"=ora_up,"oragood_dn"=ora_dn)
  return(result)
}

```

## Compare the results of bad and good analysis

```{r,compare}

# bad=ORA analysis, a list with results of up and down regulated pathways
# good=same as bad, except fixed the background problem
compare_ora_plots <- function(bad,good,libname=NULL){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  mup <- merge(badup,goodup,by=0)
  MAX=max(c(mup$es.x,mup$es.y))
  UPHEADER=paste(libname,"up")
  DNHEADER=paste(libname,"dn")

  par(mfrow=c(2,2))

  plot(mup$es.x, mup$es.y,
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original ES", ylab="corrected ES", main=UPHEADER)
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  MAX=max(-log10(c(mup$pvalue.x,mup$pvalue.y)))

  plot(-log10(mup$pvalue.x), -log10(mup$pvalue.y),
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original -log10 p-value", ylab="corrected -log10 p-value", main=UPHEADER)
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  mdn <- merge(baddn,gooddn,by=0)
  MAX=max(c(mdn$es.x,mdn$es.y))

  plot(mdn$es.x, mdn$es.y,
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original ES", ylab="corrected ES", main=DNHEADER)
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)

  MAX=max(-log10(c(mdn$pvalue.x,mdn$pvalue.y)))

  plot(-log10(mdn$pvalue.x), -log10(mdn$pvalue.y),
    xlim=c(0,MAX), ylim=c(0,MAX),
    xlab="original -log10 p-value", ylab="corrected -log10 p-value", main=DNHEADER)
  grid()
  abline(a = 0, b = 1,lwd=2,lty=2)
}

compare_ora <- function(bad,good){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  mup <- merge(badup,goodup,by=0)
  mdn <- merge(baddn,gooddn,by=0)
  bup <- subset(mup,p.adjust.x<0.05)$Row.names
  gup <- subset(mup,p.adjust.y<0.05)$Row.names
  bdn <- subset(mdn,p.adjust.x<0.05)$Row.names
  gdn <- subset(mdn,p.adjust.y<0.05)$Row.names
  sig_bad <- unique(c(bup,bdn))
  sig_good <- unique(c(gup,gdn))
  nsig_bad <- length(sig_bad)
  nsig_good <- length(sig_good)
  jac <- length(intersect(sig_bad,sig_good)) / length(union(sig_bad,sig_good))
  result=c("nsig_bad"=nsig_bad,"nsig_good"=nsig_good,"Jaccard"=jac)
  return(result)
}

```

## Run some analyses

```{r,run1}

params <- expand.grid(1:length(d),1:length(gs))
params2 <- expand.grid(names(d),names(gs))

colnames(params) <- c("d","gs")

ora_dat <- mclapply(1:nrow(params), function(i) {
  j=params[i,1]
  k=params[i,2]
  bad <- orabad(d[[j]],gs[[k]])
  good <- oragood(d[[j]],gs[[k]])
  res <- compare_ora(bad,good)
  return(res)
} , mc.cores=8)

ora_dat <- do.call(rbind,ora_dat)

res1 <- cbind(params2,ora_dat)

res1 %>%
  kbl(row.names = TRUE, caption="all results") %>%
  kable_paper("hover", full_width = F)

summary(res1$nsig_bad)
summary(res1$nsig_good)
summary(res1$Jaccard)

# NUMBER OF SIGNIFICANT SETS
ns1 <- matrix( res1$nsig_bad , ncol=9 )
rownames(ns1) <- names(d)
colnames(ns1) <- names(gs)

ns1 %>%
  kbl(row.names = TRUE, caption="significant genesets in original analysis") %>%
  kable_paper("hover", full_width = F)

ns1 <- apply(ns1,2,function(x) { x <- x[!is.na(x)] ; x <- x[is.finite(x)] ;  mean(x) } )

ns2 <- matrix( res1$nsig_good , ncol=9 )
rownames(ns2) <- names(d)
colnames(ns2) <- names(gs)

ns2 %>%
  kbl(row.names = TRUE, caption="significant genesets in corrected analysis") %>%
  kable_paper("hover", full_width = F)

ns2 <- apply(ns2,2,function(x) { x <- x[!is.na(x)] ; x <- x[is.finite(x)] ;  mean(x) } )
ns2

df <- data.frame(ns1,ns2)
df <- df[order(df$ns1),]

par(mar=c(c(5.1, 7.1, 4.1, 2.1) ))
barplot(t(df),beside=TRUE,horiz=TRUE,las=1,xlim=c(0,1250),
  xlab="no. significant sets",
  col=c("gray30","gray80"))
text(df$ns1+50,((1:nrow(df))*3)-1.5,labels=signif(df[,1],3))
text(df$ns2+50,((1:nrow(df))*3)-0.5,labels=signif(df[,2],3))
#abline(v=seq(100,1000,100),lty=2,lwd=0.5,col="gray")
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("gray80","gray30"), horiz=FALSE, cex=1.1)

mylabels <- gsub("$","%",gsub("^","+",signif(((df$ns2/df$ns1)-1)*100,3)))
text(df$ns2+160,((1:nrow(df))*3)-1,labels=mylabels,cex=1.1)

png("fig1_ORAsig.png")
par(mar=c(c(5.1, 7.1, 2.1, 2.1) ))
barplot(t(df),beside=TRUE,horiz=TRUE,las=1,xlim=c(0,1250),
  xlab="no. significant sets",
  col=c("gray30","gray80"))
text(df$ns1+50,((1:nrow(df))*3)-1.5,labels=signif(df[,1],3))
text(df$ns2+50,((1:nrow(df))*3)-0.5,labels=signif(df[,2],3))
#abline(v=seq(100,1000,100),lty=2,lwd=0.5,col="gray")
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("gray80","gray30"), horiz=FALSE, cex=1.1)

mylabels <- gsub("$","%",gsub("^","+",signif(((df$ns2/df$ns1)-1)*100,3)))
text(df$ns2+160,((1:nrow(df))*3)-1,labels=mylabels,cex=1.1)
dev.off()

pdf("fig1_ORAsig.pdf",width=7,height=7)
par(mar=c(c(5.1, 7.1, 2.1, 2.1) ))
barplot(t(df),beside=TRUE,horiz=TRUE,las=1,xlim=c(0,1250),
  xlab="no. significant sets",
  col=c("gray30","gray80"))
text(df$ns1+50,((1:nrow(df))*3)-1.5,labels=signif(df[,1],3))
text(df$ns2+50,((1:nrow(df))*3)-0.5,labels=signif(df[,2],3))
#abline(v=seq(100,1000,100),lty=2,lwd=0.5,col="gray")
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("gray80","gray30"), horiz=FALSE, cex=1.1)

mylabels <- gsub("$","%",gsub("^","+",signif(((df$ns2/df$ns1)-1)*100,3)))
text(df$ns2+160,((1:nrow(df))*3)-1,labels=mylabels,cex=1.1)
dev.off()

# JACCARD
jac <- matrix( res1$Jaccard , ncol=9 )
rownames(jac) <- names(d)
colnames(jac) <- names(gs)

signif(jac,3) %>%
  kbl(row.names = TRUE, caption="Jaccard similarity Corrected:Original") %>%
  kable_paper("hover", full_width = F)

jac2 <- apply(jac,2,function(x) { x <- x[!is.na(x)] ; x <- x[is.finite(x)] ;  mean(x) } )
jac2

par(mar=c(c(5.1, 7.1, 2.1, 2.1) ))
barplot(sort(jac2),horiz=TRUE,las=1,xlim=c(0,1), xlab="Jaccard index")

par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)

png("fig1_ORAjac.png")
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
dev.off()

pdf("fig1_ORAjac.pdf",width=7, height=7)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
dev.off()

# with SD
mysd <- apply(jac,2,function(x) { sd(x,na.rm=TRUE) } )
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
arrows(x0=cov,x1=cov,y0=jac2+mysd,y1=jac2-mysd,code=3, angle=90, length=0.1 , col="gray40")

png("fig1_ORAjacsd.png")
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
arrows(x0=cov,x1=cov,y0=jac2+mysd,y1=jac2-mysd,code=3, angle=90, length=0.1 , col="gray40")
dev.off()

pdf("fig1_ORAjacsd.pdf",width=7, height=7)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
arrows(x0=cov,x1=cov,y0=jac2+mysd,y1=jac2-mysd,code=3, angle=90, length=0.1 , col="gray40")
dev.off()

```

## Characterise sets specific to original and uncorrected analysis

```{r,spec}

compare_ora2 <- function(bad,good){
  badup <- bad$orabad_up
  baddn <- bad$orabad_dn
  goodup <- good$oragood_up
  gooddn <- good$oragood_dn
  mup <- merge(badup,goodup,by=0)
  mdn <- merge(baddn,gooddn,by=0)
  bup <- subset(mup,p.adjust.x<0.05)$Row.names
  gup <- subset(mup,p.adjust.y<0.05)$Row.names
  bdn <- subset(mdn,p.adjust.x<0.05)$Row.names
  gdn <- subset(mdn,p.adjust.y<0.05)$Row.names
  sig_bad <- unique(c(bup,bdn))
  sig_good <- unique(c(gup,gdn))
  itx <- intersect(sig_bad,sig_good)
  spec_good <- setdiff(sig_good,sig_bad)
  result=list("intersect"=itx,"spec_good"=spec_good)
  return(result)
}

params <- expand.grid(1:length(d),1:length(gs))
params2 <- expand.grid(names(d),names(gs))

colnames(params) <- c("d","gs")

ora_dat3 <- mclapply(1:nrow(params), function(i) {
  j=params[i,1]
  k=params[i,2]
  bad <- orabad(d[[j]],gs[[k]])
  good <- oragood(d[[j]],gs[[k]])
  res <- compare_ora2(bad,good)
  return(res)
} , mc.cores=8)

gslengths <- unlist(lapply(gs,function(x) { lapply(x,length) } ) )
names(gslengths) <- sapply((strsplit(names(gslengths),"\\.")),"[[",2)

speclen <- lapply(ora_dat3, function(j) {
  itxlen <- mean(gslengths[which(names(gslengths) %in% j[[1]])])
  speclen <- mean(gslengths[which(names(gslengths) %in% j[[2]])])
  return(c("itxlen"=itxlen,"spec"=speclen))
})

itxlenmx <- matrix(sapply(speclen,"[[",1),nrow=7)
speclenmx <- matrix(sapply(speclen,"[[",2),nrow=7)

rowMeans(itxlenmx,na.rm=TRUE)

rowMeans(speclenmx,na.rm=TRUE)

( rowMeans(speclenmx,na.rm=TRUE) - rowMeans(itxlenmx,na.rm=TRUE) ) / rowMeans(itxlenmx,na.rm=TRUE)

mean( ( rowMeans(speclenmx,na.rm=TRUE) - rowMeans(itxlenmx,na.rm=TRUE) ) / rowMeans(itxlenmx,na.rm=TRUE) )

```

Overall, the size of gene sets specific to the "fixed" analysis were 9.3% smaller than
the intersect.

## Make beeswarm chart

```{r,beeswarms}

gslibs <- unique(params2$Var2)

nsl <- lapply(gslibs,function(g) {
  orig <- subset(res1,Var2==g)$nsig_bad
  corr <- subset(res1,Var2==g)$nsig_good
  df1 <- data.frame("gsl"=g,"type"="orig","count"=orig)
  df2 <- data.frame("gsl"=g,"type"="corr","count"=corr)
  return(rbind(df1,df2))
} )

nsl <- do.call(rbind,nsl)
nsl$set <- paste(nsl$gsl,nsl$type)

unsl <- unstack(nsl[,c(3,4)])

#rearrange KEGG, Hallmark, HPO, Wikipath, miR, TFT, Reactome, Cellmarkers, GO
unsl <- unsl[,c(10,9, 6,5, 8,7, 18,17, 12,11, 16,15, 14,13, 2,1, 4,3)]

par(mar = c(5.1, 8.1, 4.1, 2.1))
library(beeswarm)
boxplot(unsl,horizontal=TRUE,las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05")
beeswarm(unsl,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

unsl1 <- unsl+1
boxplot(unsl1,horizontal=TRUE,log="x",las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(unsl1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)

png("fig1_ORAsigbox.png")
boxplot(unsl1,horizontal=TRUE,log="x",las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(unsl1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)
dev.off()

pdf("fig1_ORAsigbox.pdf",width=7, height=7)
par(mar = c(5.1, 8.1, 4.1, 2.1))
boxplot(unsl1,horizontal=TRUE,log="x",las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(unsl1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)
dev.off()

par(mar = c(5.1, 4.1, 4.1, 2.1))

```

## Make charts showing changes in pvalues

For bulk RNA-seq data set #1, run enrichment with all the gene set libraries.

```{r,charts1}

names(gs)

lapply(1:length(gs), function(i) {
  bad <- orabad(d[[1]],gs[[i]])
  good <- oragood(d[[1]],gs[[i]])
  compare_ora_plots(bad,good,libname=names(gs)[i])
} )

```

## Table for reviewer

```{r,table}

library(reshape2)
res2a <- as.matrix(acast(res1, Var1~Var2, value.var="nsig_bad"))
rownames(res2a) <- gsub("bulkrna","d",rownames(res2a))
rownames(res2a) <- gsub(".Rds"," original",rownames(res2a))

res2b <- as.matrix(acast(res1, Var1~Var2, value.var="nsig_good"))
rownames(res2b) <- gsub("bulkrna","d",rownames(res2b))
rownames(res2b) <- gsub(".Rds"," corrected",rownames(res2b))

res2 <- rbind(res2a,res2b)
res2 <- res2[order(rownames(res2)),]
res2c <- rowMeans(res2)
res2 <- cbind(res2,res2c)

res2 %>%
  kbl(row.names = TRUE, caption="Effect of background problem correction across seven studies and nine gene set libraries") %>%
  kable_paper("hover", full_width = F)

saveRDS(res2,"bgprob.rds")

rowMeans( ( res2b - res2a ) / res2a ) * 100

```

## Session information

For reproducibility

```{r,session}

sessionInfo()

```