### 2024-09-13 analysis/GSE158420_rna.Rmd
nice violin plot for downsampling

---
title: "GSE158420: RNA expression downsampling"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/background/

## Introduction

Here, we look at the sensitivity of pathway analysis of RNA-seq data.

```{r,packages}

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("HGNChelper")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("gplots")
  library("gridExtra")
  library("png")
  library("fgsea")
  library("parallel")
  library("RhpcBLASctl")
  library("tictoc")
  library("edgeR")
  library("vioplot")
})

CORES=16

```

## Load pathway data

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs <- gmt_import("../gmt/c2.cp.reactome.v2023.2.Hs.symbols.gmt")

names(gs) <- gsub("REACTOME_","",names(gs))
names(gs) <- gsub("_"," ",names(gs))

```

## Load RNA expression

From GEO https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158420
GSE158420_counts.txt.gz

```{r,deseq1}

if (!file.exists("GSE158420_counts.txt.gz")) {
  download.file("https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE158420&format=file&file=GSE158420%5Fcounts%2Etxt%2Egz",
    destfile="GSE158420_counts.txt.gz")
}

x <- read.table("GSE158420_counts.txt.gz",row.names=NULL)
dim(x)
x <- aggregate(. ~ row.names,x,sum)
dim(x)

#RUFKM
head(x)

```

There are some gene name errors, so I will use HGNChelper to repair them.

```{r,gene_name_fix}

human=x$row.names
new.hgnc.table <- readRDS("new.hgnc.table.rds")
fix <- checkGeneSymbols(human,map=new.hgnc.table)
x$gene <- fix$Suggested.Symbol
x$row.names=NULL
x <- aggregate(. ~ gene,x,sum)
dim(x)
row.names(x)=x$gene
x$gene=NULL
head(x)

```

MDS plot

```{r,mds}

patient <- head(as.character(unlist(lapply(1:ncol(x),function(i) {c(i,i)}))),ncol(x))
tumor <- rep(c(0,1),ncol(x)/2)
ss <- data.frame(patient,tumor)
rownames(ss) <- colnames(x)

mds <- cmdscale(dist(t(scale(x))))

cols <- gsub("1","pink",gsub("0","lightblue",tumor))

plot(mds, xlab="Coordinate 1", ylab="Coordinate 2", pch=16,col=cols,cex=2)

text(mds, labels=colnames(x),cex=0.7)

```

Now do DESeq2

```{r,deseq2}
RhpcBLASctl::blas_set_num_threads(4) #set automatic cores to 1

dds <- DESeqDataSetFromMatrix(countData = x , colData=ss, design = ~ patient + tumor)
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),assay(vsd))
dge <- as.data.frame(zz[order(zz$pvalue),])

sig <- subset(dge,padj<0.05)
SIG = nrow(sig)
DN = nrow(subset(sig,log2FoldChange<0))
UP = nrow(subset(sig,log2FoldChange>0))
HEADER = paste("mRNA", SIG , "DEGs,", UP ,"upregulated,", DN, "downregulated")
# smear
plot(log2(dge$baseMean),dge$log2FoldChange,cex=0.6,cex.axis=1.2,cex.lab=1.3,
  xlab="log2 base mean",
  ylab="log2 fold change" ,pch=19,col="#838383")
points(log2(sig$baseMean),sig$log2FoldChange,cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
top <- head(sig,20)
# volcano
plot(dge$log2FoldChange, -log2(dge$pvalue) ,cex=0.6, cex.lab=1.3,cex.axis=1.2,
 ,xlab="log2 fold change", ylab="-log2 p-value" ,pch=19,col="#838383")
points(sig$log2FoldChange, -log2(sig$pvalue),cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
# top N gene heatmap
colCols <- rep(c(0,1),ncol(x)/2)
colCols <- gsub("0","gray",colCols)
colCols <- gsub("1","black",colCols)

colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(  as.matrix(dge[1:50,c(7:ncol(dge))]), col=colfunc(25),scale="row",
  ColSideColors=colCols ,
  trace="none",margins = c(6,10), cexRow=.6, cexCol=.5, main="Top 50 genes")

```

RNA expression pathway enrichment with mitch

```{r,mitch}

rna <- mitch_import(dge, DEtype="deseq2")

res <- mitch_calc(rna, genesets=gs, priority="effect",cores=16, minsetsize=5)

mres <- res$enrichment_result
msig <- subset(mres,p.adjustANOVA<0.05)
nrow(msig)

head(subset(mres,p.adjustANOVA<0.05 & s.dist>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(mres,p.adjustANOVA<0.05 & s.dist<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

```

FGSEA

```{r,fgsea}

dge$stat[which(is.na(dge$stat))] <- 1

stat <- dge$stat
names(stat) <- rownames(dge)

pll <- lapply(1:32,function(i) {
  tic()
  fres <- fgsea(pathways=gs,stats=stat,minSize=5, nproc=i)
  tm <- toc()
  tm$callback_msg
})

fres <- fgsea(pathways=gs,stats=stat,minSize=5, nproc=8)

pll <- unlist(pll)
pll <- as.numeric(sapply(strsplit(pll," "),"[[",1))

names(pll) <- 1:32
barplot(pll)
plot(1:32,pll,type="b",pch=19)

fsig <- subset(fres,padj<0.05)
fres <- fres[order(-abs(fres$ES)),]

head(subset(fres,padj<0.05 & ES>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(fres,padj<0.05 & ES<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

fsets <- paste(fsig$pathway,as.character(fsig$ES > 0))

str(fsets)

```

FORA

```{r,fora1}

sigup <- rownames(subset(dge,padj<0.05 & log2FoldChange>0))
sigdn <- rownames(subset(dge,padj<0.05 & log2FoldChange<0))
bg <- rownames(dge)

oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
oresup$es <- (oresup$overlap / length(sigup) ) / (oresup$size / length(bg))
head(oresup,10)

oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
oresdn$es <- (oresdn$overlap / length(sigdn) ) / (oresdn$size / length(bg))
head(oresup,10)

oresup <- subset(oresup,padj<0.05)
oresup <- paste(oresup$pathway,"TRUE")

oresdn <- subset(oresdn,padj<0.05)
oresdn <- paste(oresdn$pathway,"FALSE")

osets <- c(oresup,oresdn)
str(osets)

```

Assess the effect of gene number on pathway results.

```{r,fora2}

sizes <- c(100,200,500,1000,2000,3000,4000,5000,6000,7000,8000)

ores <- lapply(sizes, function(size) {
  sigup <- head(rownames(subset(dge,padj<0.05 & log2FoldChange>0)),size)
  sigdn <- head(rownames(subset(dge,padj<0.05 & log2FoldChange<0)),size)
  bg <- rownames(dge)

  oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
  oresup$es <- (oresup$overlap / length(sigup) ) / (oresup$size / length(bg))
  nup <- nrow(subset(oresup,padj<0.05))

  oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
  oresdn$es <- (oresdn$overlap / length(sigdn) ) / (oresdn$size / length(bg))
  ndn <- nrow(subset(oresdn,padj<0.05))

  return(c(nup,ndn))
})

ores <- do.call(rbind,ores)
rownames(ores) <- sizes
colnames(ores) <- c("up","dn")
ores <- as.data.frame(ores)
ores$tot <- ores$up + ores$dn
tot <- ores$tot
names(tot) <- sizes
par(mar=c(mar = c(5.1, 5.1, 4.1, 2.1)))
barplot(tot,horiz=TRUE,las=1,xlab="no. pathways FDR<0.05",ylab="no. genes selected")

```

# Full counts

## Downsample FGSEA

This will run the FGSEA analysis at group sizes of 2, 3, 5, 10, 16, 22 and 30,
50 times each, with pseudorandom seeds from 100 to 5000 jumping up by 100.

```{r,fgsea_run,fig.height=10}

RhpcBLASctl::blas_set_num_threads(1) #set automatic cores to 1

SAMPLESIZES=c(2,3,5,10,15,20,25,30)
#SAMPLESIZES=c(3,6,10,22)
fgres <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100) # 50 repetitions
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    ss2 <- ss[mysample,]
    x2 <- x[,mysample]
    x2 <- x2[rowMeans(x2)>=10,]

    dds <- DESeqDataSetFromMatrix(countData = x2 , colData=ss2, design = ~ patient + tumor)
    res <- DESeq(dds)
    z <- results(res)
    vsd <- vst(dds, blind=FALSE)
    zz <-cbind(as.data.frame(z),assay(vsd))
    dge <- as.data.frame(zz[order(zz$pvalue),])

    dge$stat[which(is.na(dge$stat))] <- 1
    stat <- dge$stat
    names(stat) <- rownames(dge)
    fres <- fgsea(pathways=gs,stats=stat,minSize=5,nproc=3)
    fres <- fres[order(-abs(fres$ES)),]
    fsig <- subset(fres,padj<0.05)
    fsets2 <- paste(fsig$pathway,as.character(fsig$ES > 0))

    TOT=length(fsets2)
    TP=length(which(fsets2 %in% fsets))
    FP=length(which(!fsets2 %in% fsets))
    FN=length(which(!fsets %in% fsets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=16)
  do.call(rbind,dsres)
})

ssres <- fgres
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

pdf("sens_fgsea.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

dev.off()

ssres <- fgres

```

## FORA Downsample

```{r,fora_run,fig.height=10}

SAMPLESIZES=c(2,3,5,10,15,20,25,30)
fores <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100) # 50 repetitions
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    ss2 <- ss[mysample,]
    x2 <- x[,mysample]
    x2 <- x2[rowMeans(x2)>=10,]

    dds <- DESeqDataSetFromMatrix(countData = x2 , colData=ss2, design = ~ patient + tumor)
    res <- DESeq(dds)
    z <- results(res)
    vsd <- vst(dds, blind=FALSE)
    zz <-cbind(as.data.frame(z),assay(vsd))
    dge <- as.data.frame(zz[order(zz$pvalue),])

    sigup <- rownames(subset(dge,padj<0.05 & log2FoldChange>0))
    if (length(sigup) < 250) { sigup <- head(rownames(subset(dge, log2FoldChange>0)),250) }
    sigdn <- rownames(subset(dge,padj<0.05 & log2FoldChange<0))
    if (length(sigdn) < 250) { sigdn <- head(rownames(subset(dge, log2FoldChange<0)),250) }
    bg <- rownames(dge)

    oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
    oresup <- subset(oresup,padj<0.05)
    oresup <- paste(oresup$pathway,"TRUE")

    oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
    oresdn <- subset(oresdn,padj<0.05)
    oresdn <- paste(oresdn$pathway,"FALSE")

    osets2 <- c(oresup,oresdn)

    TOT=length(osets2)
    TP=length(which(osets2 %in% osets))
    FP=length(which(!osets2 %in% osets))
    FN=length(which(!osets %in% osets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=16)
  do.call(rbind,dsres)
})

ssres <- fores

par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

pdf("sens_fora.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

dev.off()

```

## Summarise downnsampling

```{r,downsample,fig.width=10}

par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

df

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
mtext("no. consistent pathways FDR<0.05")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

df

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
mtext("no. inconsistent pathways FDR<0.05")
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

df

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
mtext("Proportion of inconsistent pathways")
grid()

### PDF
pdf("sens.pdf",height=4,width=8)
par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", main="no. consistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",main="no. inconsistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",main="Proportion of inconsistent pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()
dev.off()

### PNG
png("fig4.png", width=7,height=5,units="in",res=150,pointsize=12)
par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", main="no. consistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",main="no. inconsistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",main="Proportion of inconsistent pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()
dev.off()

```

# Thin Counts

```{r,thincounts}

x2 <- thinCounts(x, prob=NULL, target.size=10000000)
colSums(x2)
x <- x2

```

## Downsample FGSEA

This will run the FGSEA analysis at group sizes of 2, 3, 6, 10, 16, 22 and 30,
50 times each, with pseudorandom seeds from 100 to 5000 jumping up by 100.

```{r,fgsea_run2,fig.height=10}

RhpcBLASctl::blas_set_num_threads(1) #set automatic cores to 1

SAMPLESIZES=c(2,3,5,10,15,20,25,30)
#SAMPLESIZES=c(3,6,10,22)
fgres <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100) # 50 repetitions
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    ss2 <- ss[mysample,]
    x2 <- x[,mysample]
    x2 <- x2[rowMeans(x2)>=10,]

    dds <- DESeqDataSetFromMatrix(countData = x2 , colData=ss2, design = ~ patient + tumor)
    res <- DESeq(dds)
    z <- results(res)
    vsd <- vst(dds, blind=FALSE)
    zz <-cbind(as.data.frame(z),assay(vsd))
    dge <- as.data.frame(zz[order(zz$pvalue),])

    dge$stat[which(is.na(dge$stat))] <- 1
    stat <- dge$stat
    names(stat) <- rownames(dge)
    fres <- fgsea(pathways=gs,stats=stat,minSize=5,nproc=3)
    fres <- fres[order(-abs(fres$ES)),]
    fsig <- subset(fres,padj<0.05)
    fsets2 <- paste(fsig$pathway,as.character(fsig$ES > 0))

    TOT=length(fsets2)
    TP=length(which(fsets2 %in% fsets))
    FP=length(which(!fsets2 %in% fsets))
    FN=length(which(!fsets %in% fsets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=16)
  do.call(rbind,dsres)
})

ssres <- fgres
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

pdf("sens_fgsea_thin.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

dev.off()

ssres <- fgres

```

## FORA Downsample

```{r,fora_run2,fig.height=10}

SAMPLESIZES=c(2,3,5,10,15,20,25,30)
fores <- lapply(SAMPLESIZES,function(n) {
  SEEDS <- seq(100,5000,100) # 50 repetitions
  dsres <- mclapply(SEEDS, function(SEED) {
    set.seed(SEED)
    mysample <- sample(1:37,n,replace=FALSE)*2
    mysample <- c(mysample,(mysample-1))
    mysample <- mysample[order(mysample)]
    ss2 <- ss[mysample,]
    x2 <- x[,mysample]
    x2 <- x2[rowMeans(x2)>=10,]

    dds <- DESeqDataSetFromMatrix(countData = x2 , colData=ss2, design = ~ patient + tumor)
    res <- DESeq(dds)
    z <- results(res)
    vsd <- vst(dds, blind=FALSE)
    zz <-cbind(as.data.frame(z),assay(vsd))
    dge <- as.data.frame(zz[order(zz$pvalue),])

    sigup <- rownames(subset(dge,padj<0.05 & log2FoldChange>0))
    if (length(sigup) < 250) { sigup <- head(rownames(subset(dge, log2FoldChange>0)),250) }
    sigdn <- rownames(subset(dge,padj<0.05 & log2FoldChange<0))
    if (length(sigdn) < 250) { sigdn <- head(rownames(subset(dge, log2FoldChange<0)),250) }
    bg <- rownames(dge)

    oresup <- as.data.frame(fora(pathways=gs,genes=sigup,universe=bg,minSize=5))
    oresup <- subset(oresup,padj<0.05)
    oresup <- paste(oresup$pathway,"TRUE")

    oresdn <- as.data.frame(fora(pathways=gs,genes=sigdn,universe=bg,minSize=5))
    oresdn <- subset(oresdn,padj<0.05)
    oresdn <- paste(oresdn$pathway,"FALSE")

    osets2 <- c(oresup,oresdn)

    TOT=length(osets2)
    TP=length(which(osets2 %in% osets))
    FP=length(which(!osets2 %in% osets))
    FN=length(which(!osets %in% osets2))
    RES=c("TOT"=TOT,"TP"=TP,"FP"=FP,"FN"=FN)
    RES
  },mc.cores=16)
  do.call(rbind,dsres)
})

ssres <- fores

par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
lapply(dots,summary)
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
lapply(ratio,summary)
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

pdf("sens_fora_thin.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

dev.off()

```

## Summarise downnsampling

```{r,downsample2,fig.width=10}

par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

df

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
mtext("no. consistent pathways FDR<0.05")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

df

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
mtext("no. inconsistent pathways FDR<0.05")
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

df

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
mtext("Proportion of inconsistent pathways")
grid()

### PDF
pdf("sens_thin.pdf",height=4,width=8)
par(mfrow=c(1,3))

dots <- lapply(fgres,function(x) { x[,"TP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"TP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=length(fsets)
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways", main="no. consistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")
grid()

legend("bottomright", legend=c("fgsea","fora"),
       col=c("blue","red"), lty=1, pch=19,  cex=1)

dots <- lapply(fgres,function(x) { x[,"FP"] })
fgres2 <- do.call(rbind,lapply(dots,summary))

dots <- lapply(fores,function(x) { x[,"FP"] })
fores2 <- do.call(rbind,lapply(dots,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="no. pathways",main="no. inconsistent pathways FDR<0.05",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()

dots <- lapply(fores,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fores,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fores2 <- do.call(rbind,lapply(ratio,summary))

dots <- lapply(fgres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(fgres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
fgres2 <- do.call(rbind,lapply(ratio,summary))

df <- data.frame(SAMPLESIZES,fores2[,"Median"],fgres2[,"Median"])
colnames(df) <- c("Sample size", "fora", "fgsea")

MAX=max(c(fores2[,"Median"],fgres2[,"Median"]))
plot(df$`Sample size`,df$fora,type="b",col="red",pch=19,xlab="Sample size",ylab="proportion",main="Proportion of inconsistent pathways",ylim=c(0,MAX))
points(df$`Sample size`,df$fgsea,type="b",col="blue",pch=19 )
grid()
dev.off()

```

## Make violin plot

```{r,vioplot}

length(osets)
length(fsets)

names(fores) <- names(fgres) <- as.character(c(2,3,5,10,15,20,25,30))

par(mfrow=c(1,3))

vioplot(lapply(fores,function(mx) { mx[,2] } ) , side="left" , col="white" , border = "red",
  plotCentre = "line", ylim=c(0,457), xlab="Sample size", ylab="no. pathways")
  mtext("no. consistent pathways FDR<0.05" )
vioplot(lapply(fgres,function(mx) { mx[,2] } ) , side="right" , col="white", border="blue" ,
   add = TRUE, plotCentre = "line")

abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")

vioplot(lapply(fores,function(mx) { mx[,3] } ) , side="left" , col="white" , border="red",
  plotCentre = "line", ylim=c(0,210), xlab="Sample size", ylab="no. pathways")
  mtext("no. inconsistent pathways FDR<0.05" )
vioplot(lapply(fgres,function(mx) { mx[,3] } ) , side="right" , col="white", border="blue",
  add = TRUE , plotCentre = "line")

fg1 <- lapply(fgres,function(mx) { TP <- mx[,2] ; FP <- mx[,3]  ; FP / (TP + FP) } )
fo1 <- lapply(fores,function(mx) { TP <- mx[,2] ; FP <- mx[,3]  ; FP / (TP + FP) } )

vioplot(fo1 , side="left" , col="white", border="red", plotCentre = "line",
  ylim=c(0,0.5), xlab="Sample size", ylab="no. pathways")
  mtext("proportion inconsistent pathways" )
vioplot(fg1 , side="right" , col="white", border="blue", add = TRUE , plotCentre = "line")

### PDF CHART
pdf("sens_vio.pdf",height=4,width=8)
par(mfrow=c(1,3))
vioplot(lapply(fores,function(mx) { mx[,2] } ) , side="left" , col="white" , border = "red",
  plotCentre = "line", ylim=c(0,457), xlab="Sample size", ylab="no. pathways")
  mtext("no. consistent pathways FDR<0.05" ,cex=0.8)
vioplot(lapply(fgres,function(mx) { mx[,2] } ) , side="right" , col="white", border="blue" ,
   add = TRUE, plotCentre = "line")

abline(h=length(osets),col="red")
abline(h=length(fsets),col="blue")

vioplot(lapply(fores,function(mx) { mx[,3] } ) , side="left" , col="white" , border="red",
  plotCentre = "line", ylim=c(0,210), xlab="Sample size", ylab="no. pathways")
  mtext("no. inconsistent pathways FDR<0.05" , cex=0.8)
vioplot(lapply(fgres,function(mx) { mx[,3] } ) , side="right" , col="white", border="blue",
  add = TRUE , plotCentre = "line")

fg1 <- lapply(fgres,function(mx) { TP <- mx[,2] ; FP <- mx[,3]  ; FP / (TP + FP) } )
fo1 <- lapply(fores,function(mx) { TP <- mx[,2] ; FP <- mx[,3]  ; FP / (TP + FP) } )

vioplot(fo1 , side="left" , col="white", border="red", plotCentre = "line",
  ylim=c(0,0.5), xlab="Sample size", ylab="no. pathways")
  mtext("proportion inconsistent pathways",, cex=0.8)
vioplot(fg1 , side="right" , col="white", border="blue", add = TRUE , plotCentre = "line")
dev.off()

```

## Session information

```{r,save}

save.image("GSE158422_sensitivity.Rdata")

sessionInfo()

```