### 2024-09-16 manuscript/manuscript.Rmd
ms work

---
title: "Two subtle problems with over-representation analysis"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: bmc-bioinformatics.csl
---

[Mark Ziemann<sup>1,2</sup>*](https://orcid.org/0000-0002-7688-6974),
[Anusuiya Bora<sup>1,2</sup>](https://orcid.org/0009-0006-2908-1352)

**Affiliations**

1. Burnet Institute, Melbourne, Australia.

2. Deakin University, School of Life and Environmental Sciences, Geelong, Australia.

(*) Corresponding author: mark.ziemann@burnet.edu.au Burnet Institute, 85 Commercial Rd Melbourne VIC 3004, Australia.

## Abstract

**Background**: Over-representation analysis (ORA) is used widely to assess the enrichment of functional
categories in a gene list compared to a background list.
ORA is therefore a critical method in the interpretation of 'omics data, relating gene lists
to biological functions and themes.
Although ORA is hugely popular, we and others have noticed two potentially undesired
behaviours of some ORA tools.
The first one we call the "background problem", because it involves the software eliminating
large numbers of genes from the background list if they are not annotated as belonging to any
category.
The second one we call the "false discovery rate problem", because some tools underestimate
the true number of parallel tests conducted.

**Results**: Here we demonstrate the impact of these issues on several real RNA-seq datasets and use
simulated RNA-seq data to quantify the impact of these problems.
We show that the severity of these problems depends on the gene set library, the number of
genes in the list, and the degree of noise in the dataset.

**Conclusion**: These problems can be mitigated by changing packages/websites for ORA or by changing to
another approach such as functional class scoring.

## Keywords

* Bioinformatics

* Pathway analysis

* Gene set enrichment analysis

* Research rigour

## Background

```{r,libs,echo=FALSE}

library("kableExtra")

```

Over-representation analysis (ORA) is a type of functional enrichment analysis (FEA) that
involves the summarisation of omics data to reflect biological differences.
FEA has become one of the most popular methods in bioinformatics,
collectively accumulating 131,332 citations as of late 2019 [@Xie2021-nv].

ORA involves the selection of genes of interest, followed by a test to ascertain whether
certain functional categories are over-represented in the selected genes.
Since its initial development in 1999 [@Tavazoie1999-bu], ORA has proliferated widely,
becoming part of many bioinformatics websites and software packages.
The most popular ORA website is DAVID [@Dennis2003-ar;@Huang2009-gi;@Sherman2022-jn],
followed by Ingenuity Pathway Analysis, a commercial software package provided by QIAGEN Inc.
ORA has been implemented into R/Bioconductor packages including `clusterProfiler`
[@Yu2012-jj;@Wu2021-wy], `limma` [@Ritchie2015-oz] and `GOseq` [@Young2010-iw].

Another popular approach to FEA is functional class scoring (FCS).
In FCS, all detected genes are ranked by their degree of differential expression followed
by a statistical test for enrichment of gene categories at either extreme of the
ranked list [@Subramanian2005-no].

According to a study that used simulated differential expression data, FCS accuracy was
superior to ORA over a range of experimental designs [@Kaspi2020-rn].
A study of 1366 PubMed Central articles featuring enrichment analysis published in 2019
indicates that seven of the top eight most popular tools were based on ORA (according to the
supplementary data)[@Wijesooriya2022-fw].
The discrepancy between performance and popularity is likely due to the relative ease of
conducting ORA as compared to FCS.
A minimal ORA might involve pasting a list of gene symbols into a text box on a website,
with results appearing nearly instantly.
On the other hand, FCS typically involves installing specific software and dealing with a
dataset representing every detected gene (~20,000 rows) and ensuring that the file
format from upstream tools is compatible with FCS packages.

Despite the popularity of ORA, there are concerns that this technique is being misused.
Due to technological and biological reasons, not all genes are detected in transcriptomic
studies, meaning that some genes will more readily appear as differentially expressed.
To address this bias, a background list of detected genes is required for comparison to the
list of selected genes (foreground list) [@Tilford2009-vx].
Failure to use a background gene list leads to dramatic changes in enrichment results,
rendering them unreliable [@Timmons2015-ki;@Wijesooriya2022-fw].
Unfortunately, use of an appropriate background list is reported in only a small fraction
(~4%) of peer-reviewed articles describing enrichment analysis [@Wijesooriya2022-fw].
Correction of p-values for multiple testing is also crucial in controlling the false positive
rate as enrichment analysis typically involves thousands of parallel tests [@Tilford2009-vx].
But appropriate p-value correction is reported in only ~54% of studies [@Wijesooriya2022-fw].
These issues have resulted in the development of best practices for end users and
stronger reporting standards [@Zhao2023-mx].

The focus of this work is to raise awareness to two existing problems in ORA implementations
of some popular enrichment tools.
The first problem is that genes belonging to the background list are removed from the analysis
entirely if they are not annotated as belonging to any functional categories.
This results in the background list appearing smaller than it should, leading to an
underestimation of fold enrichment scores and significance values.
In principle, this problem would differentially affect analyses involving different gene set
libraries, with smaller libraries such as Kyoto Encyclopedia of Genes and Genomes (KEGG)
[@Kanehisa2023-ni], which describes 2,727 genes affected more severely as compared to larger
libraries like Gene Ontology [@The_Gene_Ontology_Consortium2023-xp] which describes 19,428
genes (figures obtained from MSigDB June 2024 [@Liberzon2015-cs]).

The second problem is that adjusting p-values for multiple testing, also known as
false-discovery rate correction, is sometimes implemented improperly.
Determining which pathways should be subject to false discovery rate correction depends on
how a detection threshold is set.
Ideally, a pathway should be considered detected based on the presence of a predetermined
number of member genes in the background set.
However, some tools use the foreground set to define a pathway as detected.
This is problematic, as the act of trying to find an intersection between a foreground gene list
and a pathway is a test, even if there are no common genes.
This results in pathways being discarded from the analysis if no common genes are found in
the foreground, despite many genes being present in the background list.
This artificially reduces the number of tests conducted and makes FDR values appear lower
than they should, potentially raising the rate of false positives.
We expect this problem to have a more severe effect when the foreground list is small.

Here, we define the effect of these two problems on real RNA-seq based enrichment analysis
results and demonstrate how these two issues impact performance using simulated expression
data.
Finally, we classify popular tools regarding these two problems and provide general
recommendations for end users.

## Methods

### Expression data preparation

To quantify the effect of these two problems on RNA-seq based enrichment analysis, data from seven
transcriptomic studies were downloaded from DEE2 using the getDEE2 R/Bioconductor package
[@Ziemann2019-ux].
The criteria for selection included human as the species, presence in the DEE2 database,
passing most DEE2 quality control measures, a simple control-case experiment design with
two or three replicates per group, and with 50 or more differentially expressed genes
(FDR<0.05).
Raw data for these seven studies are available from NCBI Sequence Read Archive under the
accession numbers SRP128998, SRP038101, SRP037718, SRP096177, SRP097759, SRP253951 and
SRP068733 [@Felisbino2021-hc; @Lund2014-ds; @Rafehi2014-rv; @Keating2014-ab;
@Blanco-Melo2020-hh; @Rafehi2017-qn].
For each dataset, kallisto (v0.43.1) [@Bray2016-lu] transcript counts were aggregated to the
gene level.
Genes with fewer than 10 reads per sample on average were removed from downstream analysis.
Remaining genes passing this selection were included in the background gene set.
Differential expression analysis was conducted with DESeq2 v1.44.0 [@Love2014-wg], with
genes identified by their Ensembl identifiers.
Gene symbols were then fetched using biomaRt v2.60.0, based on Ensembl version 112 [@Durinck2009-dr].
Descriptive information of the seven datasets is provided in Table 1.

| SRA accession | Control accessions | Case accessions | Comparison | No. detected genes | No. genes with FDR<0.05 | Reference |
| --- | --- | --- | --- | --- | --- | --- |
| SRP128998 | SRR6467485, SRR6467486, SRR6467487 | SRR6467479, SRR6467480, SRR6467481 | Normal glucose versus high glucose | 15,635 | 3,472  | [@Felisbino2021-hc] |
| SRP038101 | SRR1171523, SRR1171524, SRR1171525 | SRR1171526, SRR1171527, SRR1171528 | Control versus azacytidine treated cells | 13,926 | 3,590 | [@Lund2014-ds] |
| SRP037718 | SRR1168228, SRR1168229, SRR1168230 | SRR1168225, SRR1168226, SRR1168227 | Control versus SAHA treated cells | 15,477 | 4,988 | [@Rafehi2014-rv] |
| SRP096177 | SRR5150595, SRR5150596, SRR5150597 | SRR5150592, SRR5150593, SRR5150594 | Control versus Set7 knock-down cells | 15,607 | 5,152 | [@Keating2014-ab] |
| SRP097759 | SRR5201525, SRR5201526 | SRR5201527, SRR5201528 | GFP control vs SAHH overexpression | 19,139 | 62 | Unpublished |
| SRP253951 | SRR11517674, SRR11517675, SRR11517676 | SRR11517677, SRR11517678, SRR11517679 | Mock vs SARS-CoV-2 infection | 15,182 | 8,588 | [@Blanco-Melo2020-hh] |
| SRP068733 | SRR3112216, SRR3112217, SRR3112218 | SRR3112219, SRR3112220, SRR3112221 | Control vs EP300 knock-down | 14,255 | 7,365 | [@Rafehi2017-qn] |

: Table 1. Seven selected RNA-seq datasets for benchmarking.

### Examining the background problem using real transcriptome data

Human gene sets were obtained from MSigDB v2023.2 [@Liberzon2015-cs], and included
KEGG Medicus [@Kanehisa2023-ni],
Reactome [@Milacic2024-iv],
Wikipathways [@Agrawal2024-lr],
MicroRNA targets from miRdb [@Chen2020-wo],
transcription factor targets from GTRD [@Kolmykov2021-mn],
Gene Ontology (GO) [@The_Gene_Ontology_Consortium2023-xp],
Human Phenotype Ontology (HPO) [@Kohler2021-ha],
Cell markers and
Hallmark pathways [@Liberzon2015-cs].
To demonstrate the background problem, differentially expressed genes with
FDR<0.05 were subset, with separate lists for up and downregulated genes.
If fewer than 200 genes met this criterion, then the 200 genes with the smallest p-values
were selected, as suggested by a previous report [@Tarca2013-xh].
These gene lists were subjected to ORA using clusterProfiler's `enricher` function
v4.12.0, using a minimum set size of 5 and a no maximum set size.

To systematically assess the impact of the background problem, we conducted ORA with
clusterProfiler using the different gene set libraries described above for the seven RNA-seq
datasets.
ClusterProfiler was used with the same parameters as above, and we devised a simple workaround
to the background problem, which is to append the entire background list as a gene set to the
library.
This forces clusterProfiler to retain all background genes.
The ORA results were then filtered for significant sets using an FDR threshold of 0.05.
Jaccard index was used to compare original and corrected analyses.

### Examining the FDR problem

For this test, 2000 of the top up and down regulated genes were selected from each
differential expression dataset based on p-value.
A workaround for the FDR problem was devised by filtering the gene sets by the presence of
at least two genes in the background list prior to running clusterProfiler.
After running clusterProfiler using the parameters above, the number of gene sets in the
results was quantified, and compared to the number that should have been reported using the
two gene limit.
The difference between these numbers (*n*) represents the number of missing gene sets.
To account for these tests, *n* values of 1 were appended to the p-values obtained by
clusterProfiler, followed by FDR correction in R, to obtain the properly corrected FDR values.
As above, analysis of seven datasets was done with nine gene set libraries, the significant
sets were selected and the Jaccard statistics were collected for each run to compare original
and corrected analysis.

As we predicted that this effect could be more severe for smaller foreground lists, we
performed parallel analysis with foreground gene lists with sizes varying between 125 and
2,000 genes, as ranked by FDR value.

### Quantifying the effect of these problems on precision and recall using simulated expression data

RNA-seq counts for SRA run accession ERR2539161 were obtained from DEE2 [@Mo2018-zb].
This dataset was selected due to its high number of assigned reads (367M).
A library of 200 random gene sets, each containing 30 genes, was generated.
To generate differential expression profiles, 5% of gene sets were selected to be
differentially expressed, with equal numbers of up- and down-regulated pathways.
Pseudosamples were generated by first downsampling the counts using the `thincounts`
function of the `edgeR` package v4.2.0 [@Robinson2010-eb] to 20M reads, followed by the
addition of some extra variation drawn from a normal distribution with mean of 1 and
standard deviation varying between 0.1 and 0.6, this value selected randomly for each gene.
Three "control" and five "case" pseudosamples were generated.
For the genes selected to be differentially expressed, expression values were multiplied by a
log fold change in the case samples of 0.5 for upregulated genes and -0.5 for downregulated
genes.
Genes that were selected to be both up and downregulated were left unchanged.
With these expression profiles, DESeq2 was used for differential analysis, and the results
were subjected to clusterProfiler which suffers from both problems,
clusterProfiler with the background problem fix, clusterProfiler with the FDR problem fix,
clusterProfiler with the fixes to background and FDR problems, an ORA function called `fora`
from `fgsea` bioconductor package v1.30.0 that does not suffer these problems, and `fgsea` a
fast implementation of the preranked gene set enrichment analysis (GSEA) algorithm
[@Korotkevich2016-gd].
For the ORA functions above, significantly differentially expressed genes with FDR 0.05 were
selected for the foreground gene set.
If fewer than 200 genes were detected as significant, then 200 genes with the smallest
p-values were used for ORA, like a previous report [@Tarca2013-xh].
For `fgsea`, the DESeq2 test statistic was used for scoring differential expression.
After enrichment analysis with these approaches, the significant up and down regulated
sets were selected and compared to the ground truth to calculate precision, recall
and F1 score.
At each value of added variance, 1000 replications were conducted.
No set seed was used, but results were similar when run on different computer systems.
All analyses were undertaken with R v4.4.1 in a Docker container.

### Comparing ORA and FCS methods by downsampling real RNA-seq data

To understand the sensitivity and false positive rate of ORA and FCS methods, we employed a downsampling
approach to a previously published dataset derived from 37 lung cancer patients, having tumour and normal
tissues for each patient [@Cho2022-ri].
These RNA-seq based gene expression counts were downloaded from NCBI GEO under accession number GSE158420.
After reading them into R, we noticed some gene names were converted to dates [@Abeysooriya2021-qt], so we
used the HGNChelper package to fix them [@Oh2022-gj].
DESeq2 was then used for differential expression analysis correcting for patient-of-origin to identify genes
differentially expressed in tumour compared to normal tissue.
FCS was conducted using fgsea using the DESeq2 test statistic to score each gene.
ORA was conducted using fora, using genes with differential expression FDR<0.05.
Separate tests were conducted for up- and down-regulated genes.
For the ORA background, genes with a mean of 10 or more reads per sample on average across the comparison were
included.
Reactome pathways were used.
For downsampling, sample size was varied between 2 and 30 patients, selected pseudorandomly with a set seed,
followed by parallel analysis with fgsea and fora pipelines, using a significance filter of FDR<0.05.
We then calculated the number of pathways identified in downsampled data that were consistent with the full
dataset.
The number of consistent pathways gives an indication of sensitivity of each method.
Then, we calculated the number of pathways identified in the downsampled data that were inconsistent with the
full dataset.
The proportion of significant pathways classified as inconsistent provides an estimate of the real false
discovery rate.
This was repeated 50 times at each sample size, and the median values were reported.

## Results

### Understanding the background problem

A workaround to eliminate the background problem was developed and we used it to compare the
results of ORA of seven datasets with the original and corrected methods for nine different
gene set libraries.
The number of statistically significant results was calculated for each of the gene set
libraries before and after correcting the problem (**Figure 1A**).
Each study was affected to a similar degree (mean uplift 125% ±199%), except for #5 which saw the number of
significant GO terms increase from 8 to 54 after correction (**Table S1**).
The similarity between results, as quantified with the Jaccard similarity, was highest for
Cellmarkers and GTRD at 0.76 and 0.77 respectively, and the lowest similarity was recorded for
HPO and KEGG with 0.20 and 0.48 respectively.
Observed Jaccard scores correlated with the number of genes annotated to one or more functional
categories in the gene set library (**Figure 1B**).
This result suggests that ORA with smaller gene set libraries like HPO and KEGG are impacted
more severely.
We investigated discrepant results for data set #1 (SRP128998) downregulated genes in HPO gene sets.
A scatterplot of significance scores (**Figure 1C**) shows severe deviation in p-values in gene sets in the largest
quintile (109-1382 members detected), suggesting larger gene sets are more severely affected by the background
problem.

![**Figure 1A. Number of statistically significant pathways (FDR<0.05) in seven independent RNA-seq studies with and
without correction of the background problem.
Separate tests were conducted for up- and down-regulated gene lists.
Mean values are depicted as an “x”. The percent increase in the mean number of significant sets are shown.
“*” indicates that 1 has been added to all values before log transformation.](
../analysis/fig1_ORAsig.png "Figure 1A
shows that correcting the background bug improves the number of significant sets between 12 and
229%.")

![**Figure 1B. Impact of the background problem is worse for smaller gene set libraries.**
Mean Jaccard similarity index was calculated for original and corrected ORA for seven independent datasets.
The number of genes represented in one or more functional sets is shown on the x-axis, and the mean Jaccard index is
shown on the y-axis; whiskers represent the standard deviation.](
../analysis/fig1_ORAjac.png "Figure 1B shows that the background bug impacts smaller gene sets libraries like KEGG
and HPO more severely.")

![**Figure 1C. Significance values (-log10 p-values) for dataset #1 (SRP128998) enrichment analysis
using HPO gene sets with and without correction of the background problem.**
Gene sets were divided into quintiles based on the number of genes detected in the dataset.](
../analysis/fig1_HPOdn.png "Figure 1C shows that larger gene sets are affect the p-values more severely by the
background bug.")

### Understanding the FDR problem

A workaround for the FDR problem was developed, then original and corrected
analyses were conducted for the seven datasets and nine gene set libraries.
Results showed no major differences in the number of significant sets between
original and corrected analyses (**Figure 2A**).
The impact of the FDR problem was similar across these seven datasets (**Table S2**).

We posited that a shorter foreground gene list length might exacerbate the problem, so we performed
parallel analysis with foreground lists of length between 125 and 2,000 genes (**Figure 2B**).
A clear trend was observed for GO, Reactome, Cellmarkers and TFT GTRD, with Jaccard similarity
being significantly reduced when foreground gene lists were shorter than 500 genes.
Interestingly, the drop in Jaccard similarity was more severe for GO and Reactome sets as
compared to Cellmarkers and TFT GTRD sets and may be associated with the size of the gene
sets in the library.
Median gene set size for these are GO:18, Reactome:11, TFT GTRD:287 and Cellmarkers:115.
These results demonstrate that the FDR problem is more severe when dealing with shorter
foreground lists, and this is exacerbated when functional categories in the gene set library
are also small.

![**Figure 2A. Effect of correcting the FDR problem.**
Figures represent the mean number of significant sets with and without correction of the FDR problem over seven
independent RNA-seq datasets.
Foreground list consisted of 2000 genes with the smallest p-values.
Mean values are depicted as an “x”.
“*” indicates that 1 has been added to all values before log transformation.](
../analysis/fig2_ORAfdr.png "Figure 2A shows that correcting the FDR bug slightly reduces the number of significant
sets.")

![**Figure 2B. Impact of the FDR problem is worse for shorter foreground gene lists.**
Mean Jaccard index between original and corrected analysis were calculated for seven RNA-seq datasets with gene sets
from GO, Reactome, Cellmarkers and TFT GTRD; whiskers represent the standard deviation.](
../analysis/fig2_ORAsz.png "Figure 2B shows that the impact of the FDR bug is worse when
foreground gene lists are shorter than 500.")

### Impact of these two problems on accuracy determined with simulated data

Simulated differential expression profiles with *a priori* changes to predetermined pathway
genes were analysed in parallel by clusterProfiler default and with mitigations for
problems described above.
In addition, these data underwent ORA with fora which does not suffer from these two
problems.
For comparison, fgsea was used to compare the performance of these two ORA methods against a
prototypical FCS method.
These methods were tested with different levels of introduced variation and the
precision, recall and F1 scores were recorded (**Figure 3A**).
Mitigating the background problem improved mean recall (by 2.9%), however it also caused
a decrease in mean precision (1.3% lower).
Addressing the FDR problem improved mean precision by 11.7%, but mean recall dropped by 16.6%.
Addressing both problems improved mean precision by 9.7% but decreased mean recall by 11.7%.
Under these conditions, the FDR problem affected results more severely than the background
problem.
Results obtained for clusterProfiler after implementing the two mitigations were identical to
fora to three significant figures.
fgsea recorded mean precision of 0.939, similar to the ORA methods, but recall was far superior,
with fgsea scoring 44.9% higher than default clusterProfiler and 64.2% higher than fora.
As a result, fgsea's overall accuracy as defined by F1 index was 35.4% higher than fora and
28.9% higher than default clusterProfiler.

When looking at the accuracy profile at different levels of added noise (**Figure 3B**),
ORA techniques show high precision when the added noise is low (<0.2), but
precision drops at higher levels of added noise.
The rate of drop in precision was more severe for approaches that suffer from the FDR problem.
The precision of fgsea was relatively stable despite high levels of added noise, ranging from
0.95 to 0.91.
The recall of fgsea was superior to all ORA approaches across all levels of added noise.

![**Figure 3A. Impact of correcting these two problems on accuracy of simulated expression
data.**
(A) Precision, recall and F1 scores across simulations.
CP; clusterProfiler, CP BG fix; clusterProfiler with a workaround to the background problem,
CP FDR fix; clusterProfiler with a workaround to the FDR problem,
CD BG and FDR fix; clusterProfiler with mitigations for both background and FDR problems,
fora; ORA function of the fgsea package,
fgsea; an FCS method similar to GSEA.
Mean values are depicted as an “x”.
](../analysis/sim/fig3_bars.png "Figure 3A shows that fora is more
precise than clusterProfiler, and fgsea has higher recall than ORA methods.")

![**Figure 3B. Accuracy of clusterProfiler, fora and fgsea with simulated data with various
levels of added noise.**
Precision, recall and F1 score are shown with the addition of different amounts of random noise to expression levels.
The purple line for clusterProfiler with mitigations for both problems is not visible as the performance was
identical to fora.
](../analysis/sim/fig3_sim.png "Figure 3B shows that ORA methods have deteriorating accuracy
with noisy data as compared to fgsea, a FCS method.")

### Exploring sensitivity in real RNA-seq data

To understand whether FCS has better sensitivity than ORA in a real setting, we conducted a
systematic downsampling of a gene expression dataset consisting of normal and tumour samples
of 37 lung cancer patients.
In the full set of patient samples, fora (ORA) yielded 375 significantly differentially
regulated Reactomes, while fgsea (FCS) gave 480 (FDR<0.05).
Fgsea yielded consistently more significant pathways than fora in downsampled data
(**Figure 4A**).
Fgsea also identified a larger proportion of pathways in downsampled data compared to fora.
For example at a sample size of 5, fgsea identified 73.3% (median) of pathways from the full set as
compared to 41.5% for fora, confirming better recall for FCS over ORA.
It was noted that fgsea identified more pathways in downsampled data that were
inconsistent with the full dataset as compared to fora (**Figure 4B**).
However, inconsistent pathways were a smaller proportion of all findings for fgsea as
compared to fora (**Figure 4C**), indicating an overall lower false positive rate for FCS
compared to ORA methods.

![**Figure 4. Downsampling confirms higher sensitivity for FCS over ORA.**
(A) Violin plot showing the number of significant pathways identified in downsampled data that are consistent with
findings in the full dataset, using FCS (fgsea) and ORA (fora) methods.
Downsampling was repeated 50 times.
The blue and red horizontal lines indicate the number of significant pathways identified
in the full set of 37 patients using fgsea and fora respectively.
(B) The number of significant pathways identified in downsampled data that are not
consistent with the full dataset.
(C) The proportion of pathways in downsampled data that are inconsistent with the full
dataset.](
../analysis/fig4.png "Figure 4A shows that fgsea is more sensitive than fora. Figure 4B shows fgsea has higher
absule number of false positives, but Figure 4C shows that as a proportion of all findings fgsea has lower false
positive rate than fora.")

## Discussion

The background problem came to our attention when a doctoral student in our department
was puzzled by results obtained from clusterProfiler when using a custom set of pathways.
This exact issue was the subject of several posts to online bioinformatics forums dating to
~2020 where confused users were wondering why the number of background genes changed so much.
The effect of excluding genes without functional annotations causes the background list to appear
smaller than it would otherwise be, causing enrichment scores to appear smaller.
The results shown here indicate the background problem affects smaller gene set libraries
like KEGG and HPO more severely as compared to Cellmarkers and GO libraries which describe
a larger proportion of all genes.
ClusterProfiler maintainers recently provided an option to prevent the loss of unannotated
genes from the background `options(enrichment_force_universe = TRUE)`, however this feature
isn't yet described in the official documentation.

The FDR problem is potentially more serious in that it could lead to increasing false
positives under specific conditions; when the foreground list is short, and the median pathway
size is relatively small.
These false positives arise due to an underestimate of the actual number of tests conducted.
In extreme cases, correction of this problem can lead to around half of the statistically
significant results changing (see Figure 2B).

So end-users can explore the effect of these two problems on their work, we have developed
an R/Shiny web application (see "Availability of data and materials" below).
Users can upload their foreground and background lists, select their preferred gene set
library and compare results from default and corrected analyses.

ClusterProfiler is not the only tool to suffer from these problems, they are widespread
(**Table 2**). Still, there are some tools free of such problems.
For example, fora is a good option for R-based workflows, however the function does not calculate
the fold enrichment score,
so users will need to do it themselves [@Korotkevich2016-gd].
ShinyGO is an excellent alternative for web based analysis [@Ge2020-lu].
It doesn't suffer from the two issues described here, it is easy to use,
and has a high degree of reproducibility thanks to its stand-alone versioned docker
container option for local execution.
One problem that all of these tools listed in Table 2 have is that a background list is optional.
Sometimes this feature is hidden under "advanced options".
Given the strong biological and technical biases that are present in contemporary omics such as single cell
and spatial transcriptomics, it should be a mandatory step for users to provide a custom background list.

| Tool | Version | Type | Provides FDR values | Provides enrichment score | Proper background handling | Proper FDR | Reference |
| --- | --- | --- | --- | --- | --- | --- | --- |
| DAVID | v2023q4 | Web | Yes | Yes | No | Yes | [@Sherman2022-jn] |
| Panther | v19 | Web | Yes | Yes | Yes | Yes | [@Mi2019-ak] |
| Enrichr | June 2023 | Web | Yes | Yes | Yes | No | [@Kuleshov2016-og] |
| KOBAS-i | KOBAS 3.0 | Web | Yes | No | Yes | No | [@Bu2021-ak] |
| WebGestalt | 2024 | Web | Yes | Yes | No | Yes | [@Elizarraras2024-nr] |
| g:Profiler | Feb 2024 | Web | Yes | No | No | Yes | [@Kolberg2023-nr] |
| STRING-DB | v12.0 | Web | Yes | Yes | Yes | Yes | [@Szklarczyk2023-an] |
| ShinyGO | v0.80 | Web | Yes | Yes | Yes | Yes | [@Ge2020-lu] |
| BinGO | v3.0.5 | Cytoscape | Yes | No | No | No | [@Maere2005-gq] |
| ClueGO | v2.5.10 | Cytoscape | Yes | No | No | No | [@Bindea2009-jn] |
| clusterProfiler | v4.12.0 | R package | Yes | No | No | No | [@Wu2021-wy] |
| topGO | v2.56.0 | R package | No | No | No | Yes+ | [@Adrian_Alexa2017-nf] |
| GOseq | v1.56.0 | R package | No | No | No* | Yes+ | [@Young2010-iw] |
| goana/kegga | limma v3.60.3 | R package | No | No | No** | Yes+ | [@Ritchie2015-oz] |
| fora | fgsea v1.30.0 | R package | Yes | No^ | Yes | Yes | [@Korotkevich2016-gd] |

: Table 2. Classification of several popular freely available ORA tools regarding the provision of information such as
enrichment scores and FDR values, as well as behaviours regarding the two algorithmic issues described above.
(*) Denotes functions where options are provided to adjust this behaviour and are documented in the user manual.
(**) Documentation states unannotated genes are discarded with default settings using whole genome
background, but not when a custom background is used.
(+) Denotes that although no FDR values are provided, the results include sets where no overlaps were found, so users can adjust raw p-values themselves.
(^) Denotes enrichment scores will be included in the next Bioconductor release.

The purpose of this work was to thoroughly characterise two subtle problems with ORA, but there
are some bigger problems noted.
Firstly, all variations of ORA were less accurate (based on F1 score) as compared to FCS in this
simulation work, confirming a previous report [@Kaspi2020-rn].
Specifically, precision of ORA deteriorated with greater levels of added noise, while FCS
was relatively stable.
This means researchers can be more confident with FCS results, even if the data are
noisy.
The higher recall of FCS demonstrated in simulated differential expression profiles and confirmed
with downsampled real cancer RNA-seq data means that researchers will have more power to identify
pathways relevant to the biological processes being studied.
Secondly, many popular ORA tools listed in Table 2 do not provide enrichment scores in their
default output tables.
Enrichment scores are a surrogate for effect size, and without them, users may over-rely
on significance values when interpreting enrichment results.
Lastly, ORA results can vary dramatically depending on the number of genes selected in the
foreground, a mostly arbitrary choice.

These issues together with previously mentioned pitfalls such as lack of background
correction and methodological misreporting [@Timmons2015-ki;@Wijesooriya2022-fw], suggest
that if the data can be scored/ranked, methods like FCS are strongly recommended over ORA.
ORA is better suited to other cases where genes are classified into binary groups.
For example when investigating the over-represented categories in gene expression modules
[@Langfelder2008-hu], or with variant-harbouring genes associated with a disease [@White2019-ps].
In these cases, we would recommend using a tool from Table 2 that reports FDR values and enrichment scores,
and avoids background and FDR problems.
Doubts also arise when choosing options and parameters for ORA.
Most important is to use a background gene list that reflects the true universe of genes detected robustly in the
assay.
Secondly, significance of enrichment results must be based on the FDR-corrected p-values.
Selection criteria for foreground genes is another crucial choice; for this we endorse the approach of
Tarca et al, [@Tarca2013-xh] which accommodates cases where relatively few individual genes meet the FDR threshold.
Some ORA tools have a filter for gene set size.
We’d recommend a minimum size of 5-10 in order to avoid missing small gene sets that exhibit strong enrichments.
Interpret with caution enrichment results where the intersection of input gene list and pathway is ≤3 genes.
A maximum gene set filter is not required if end results are interpreted using enrichment scores together with FDR
values.
The choice of pathway database also influences the conclusions that can be drawn [@Mubeen2019-gw;@Karp2021-ue].
Our suggestion is to use comprehensive and open-source pathway databases such as Reactome and Gene Ontology
Biological Process.
Proprietary databases like KEGG and Ingenuity are less comprehensive, have restrictive conditions for non-academic
users and do not provide public access to historical database versions which inhibits future reproducibility.

## Declarations

### Ethics approval and consent to participate

Not applicable.

### Consent for publication

Not applicable.

### Availability of data and materials

Publicly available data were obtained from Digital Expression Explorer 2 (dee2.io) and
NCBI GEO using the accession numbers mentioned in the Methods section.

Reproduction of the results here can be achieved using the GitHub code repository GitHub
(https://github.com/markziemann/background) and Docker image
(https://hub.docker.com/r/mziemann/background), which have both been deposited to Zenodo
for long-term preservation (https://zenodo.org/record/TODOXXX).

The R/Shiny tool for comparing ORA with and without the two problems is currently
online at https://oratool.ziemann-lab.net/ and available as a
downloadable docker container for local execution
(https://hub.docker.com/repository/docker/mziemann/background_app/general).

### Competing Interests

The authors declare that they have no competing interests.

### Funding

This work was supported by Burnet Institute and Deakin University.
Authors declare that no grants were involved in supporting this work.

### Author Contributions

Conceptualization: MZ, AB.
Data Curation: MZ.
Formal Analysis: MZ.
Funding Acquisition: N/A.
Investigation: MZ, AB.
Methodology: MZ, AB.
Project Administration: MZ.
Resources: MZ.
Software: MZ, AB.
Supervision: MZ.
Validation: AB.
Visualization: MZ, AB.
Writing – Original Draft Preparation: MZ, AB.
Writing – Review & Editing: MZ, AB.

### Acknowledgements

We thank Ms Megan Soria and Ms Kaumadi Wijesooriya (Deakin University) for discussions on the concept and manuscript.
This research was supported by use of the Nectar Research Cloud, a collaborative Australian research
platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).
The authors gratefully acknowledge the contribution to this work of the
Victorian Operational Infrastructure Support Program received by the Burnet Institute.

## Bibliography